### 2019-05-27 Report/Progress_Report/paper_summaries.tex
writing up paper summaries as preperation for background report

\documentclass[12pt]{article}

\title{Paper Summaries}
\author{Peter Robertson}
\date{}

\begin{document}
\maketitle

\section{GANs Based Methods}
\subsection{Generative Adversarial Network}
The original GANs paper proposed by Ian Goodfellow \cite{Goodfellow2014}.
This paper uses two feed forward neural networks (as proof of concept) as generator and discriminator.
The original loss function is used, although it is mentioned that even at this stage issues with it are highlighted.
Argument made to say that while learning a probability distribution is great, the goal is to produce samples from it, so generating sample from an implicit distribution is as useful.
No need to go into extreme depth as this was covered in the Deep Learning lectures.

\subsection{Conditional Generative Adversarial Nets}
Conditional GANs \cite{Mirza2014} extend from Goodfellow's work \cite{Goodfellow2014} by concatenating an input to the input noise given the generator.
This is performed on the MNIST Dataset \cite{LeCun2010} by conditioning the generator on a given digit to produce.

\subsection{DCGAN}
This paper \cite{Radford2016} takes the concept of adversarial training of neural networks and applies it with convolutional neural networks on the CIFAR10 dataset \cite{Krizhevsky2009}.
The paper also attempts to make some progress in stabilizing the training of GANs models with the use of Batch Normalization before the input to each subsequent layer.
Experimentally this is shown to assist training by helping gradients to flow through the network and poor weight initialization.

\subsection{Wasserstein GANs}
In an attempt to stabilize the training of GANs models, the Wasserstein or 'Earth-Mover' loss function is proposed \cite{Arjovsky2017}.
By using the Wasserstein loss function, the generator and discriminator no longer have to alternate training epochs to try to maintain model stability.
By using the Wasserstein loss function, the discriminator (critic as it is called in this paper) is trained to an optimum point, at which time the loss provided to the generator is at a maximum, making training the generator easier.
The generator can then be trained to an optimum position, at which time, training the critic can continue again.
This method has greatly improved stability and is fairly robust to a large range of model architectures which have been tested.
The generator and critic also no longer have to have a similar capacity.

To achieve this, the Lipschtiz condition must be met.
This is achieved by clipping the weights on the model.
The paper points out that this method of achieving the Lipschtiz condition is non-ideal and is a point of further work.

Other points of note, the loss plot produced by the WGAN model seems to correlate with perceived image quality, which could not be said for precious loss functions.

\subsection{Improved Training of Wasserstein GANs}
WGAN+GP \cite{Gulrajani2017}

\subsection{Progressive Growing of GANs for Improved Quality, Stability, and Variation}
Progressive growing GANs \cite{Karras2017b}

\subsection{End-to-End Speech-Driven Facial Animation with Temporal GANs}
These guys are downstairs \cite{Vougioukas2018}

\subsection{Self-Attention Generative Adversarial Networks}
Attention GANs, Goodfellow involved \cite{Zhang2018}

\subsection{Few-Shot Adversarial Learning of Realistic Neural Talking Head Models}
This paper uses meta-learning on conjunction with adversarial learning methods \cite{Zakharov2019}.
The paper aims to learn an encoding which represents a person's face and then use this to generate video of the subject moving into an unseen position.
The model initially learns an encoding based on several frames from an input video from random positions in the video.
Once the subject encoding has been learnt, the model continues to learn to construct frames of video of the subject using facial landmarks to drive the positioning.
Training is performed with existing positions in the video which the GAN model is tested on if the presented frame is real or generated.
After training, the model is capable to producing subjects in positions unseen in the video.

In regards to 3D modelling, a similar method could be used to encode a subject in a very similar manner to what has been performed here.
With this subject encoding, a generic neutral mesh template could be replaced by a neutral subject specific mesh template, reducing the work of the generator to just having to capture realistic movement as opposed to movement and reproducing the subject.

\section{Other Generative Methods}
\subsection{Capture, Learning, and Synthesis of {3D} Speaking Styles}
\cite{Cudeiro2019}

\subsection{Audio-driven facial animation by joint end-to-end learning of pose
and emotion}
Nvidia facial animation from speech \cite{Karras2017a}

\subsection{End-to-end Learning for 3D Facial Animation from Raw Waveforms of
Speech}
\cite{Pham2017}

\subsection{Synthesising 3D Facial Motion from "In-the-Wild" Speech}
Imperial Paper \cite{Tzirakis2019}

\section{3D Modelling}
\subsection{Direct-Manipulation of Blendshapes}
This paper \cite{Lewis2010} provides a reasonable overview of what blendshapes are.
It's a slightly older paper from 2010, but the techniques in the field of animation haven't changed.
Animating one second of video may take a skilled animator an hour using blendshape models.
Blendshapes are linear weighted sum of blendshape targets.
These targets can be facial expressions, or approximations of facial muscles.
A blendshape can have 100+ degrees of freedom, while facial meshes have many many more.
Limiting the degrees of freedom to a human animator to 100 is still huge, but this may oversimplify the problem to a computer.
Facial mesh may have 5000-10000 vertices, perhaps using PCA to reduce this down is smart.
Or maybe using a technique similar to Nvidia's Progressively growing GAN \cite{Karras2017b} may allow the DOF to be slowly increased.

\subsection{Learning a model of facial shape and expression from 4D scans}
This paper creates the FLAME head model which has been designed for animation \cite{Li2017}.
With the aim of being computationally efficient and compatible with game engines, this model only contains 5023 vertices.
Aim is to be a good middle ground between current simple models which are unrealistic and high end models which are very time consuming.
Model uses very large dataset made up of multiple datasets.
Current comparison is FaceWarehouse
Does some work to address the issue of registering 4D scans.

\section{Lip Reading Papers}
\subsection{Out of Time: Automated Lip Sync in the Wild}
Lip sync in the wild \cite{Chung2016}

\section{Speech Recognition}
\subsection{Deep Speech: Scaling up end-to-end speech recognition}
Mozilla Deep Speech \cite{Hannun2014}

\section{Attention Based Mechanism}
\subsection{Attention is All you Need}
Google Attention paper \cite{Vaswani2017}

\subsection{Show, Attend and Tell: Neural Image Caption Generation with Visual
Attention}
Attention applied to images \cite{Xu2015}

\bibliographystyle{unsrt}
\bibliography{../ref}
\end{document}

### 2019-05-27 Report/Progress_Report/paper_summaries.tex
adding comments to WGAN, WGAN+GP

@@
 \subsection{Improved Training of Wasserstein GANs}
-WGAN+GP \cite{Gulrajani2017}
+To improve the Lipschitz constraint condition to the WGANs \cite{Arjovsky2017}, an alternate method is proposed by using gradient penalty \cite{Gulrajani2017}.
+Paper examines the distribution of weights with WGAN and found that weights are bunched around clipping limits.
+With WGAN+GP the weights form a normal distribution, allowing more complex functions to be captured as clipped weights for a bias to simple functions, limiting the model capacity.
+No batch normalization is used, as this is no longer valid when using gradient penalty.
+This loss function seems to be able to train any architecture, even ResNet architectures which had currently been unsuccessful when used in GANs models.
+Training speed is also improved compared with WGAN.
 

### 2019-05-30 Report/Progress_Report/Background.tex
renamed background report to fit CATE spec

\documentclass[12pt,twoside]{report}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Definitions for the title page

\newcommand{\reporttitle}{Progress Report}
\newcommand{\reportauthor}{Peter Robertson}
\newcommand{\supervisor}{Dr Stefanos Zafeiriou}
\newcommand{\degreetype}{Computing Science / Machine Learning}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% load some definitions and default packages
\input{includes}

% load some macros
\input{notation}

\date{September 2019}

\begin{document}

% load title page
\input{titlepage}

% page numbering etc.
\pagenumbering{roman}
\clearpage{\pagestyle{empty}\cleardoublepage}
\setcounter{page}{1}
\pagestyle{fancy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%--- table of contents
\fancyhead[RE,LO]{\sffamily {Table of Contents}}
\tableofcontents

\clearpage{\pagestyle{empty}\cleardoublepage}
\pagenumbering{arabic}
\setcounter{page}{1}
\fancyhead[LE,RO]{\slshape \rightmark}
\fancyhead[LO,RE]{\slshape \leftmark}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}

\begin{figure}[tb]
\centering
\includegraphics[width = 0.4\hsize]{./figures/imperial}
\caption{Imperial College Logo. It's nice blue, and the font is quite stylish. But you can choose a different one if you don't like it.}
\label{fig:logo}
\end{figure}

Figure~\ref{fig:logo} is an example of a figure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Background}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Contribution}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Experimental Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusion}

%% bibliography
\bibliographystyle{apa}

\end{document}

### 2019-05-30 Report/Progress_Report/Background.tex
Bullet point plan for progress report. Need to read into lip reading papers right now

@@
-\documentclass[12pt,twoside]{report}
+\documentclass[12pt]{article}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{geometry}
+ \geometry
+ {
+     a4paper,
+     left=25mm,
+     right=25mm,
+     top=25mm,
+     bottom=25mm,
+ }
 
-% Definitions for the title page
-
-\newcommand{\reporttitle}{Progress Report}
-\newcommand{\reportauthor}{Peter Robertson}
-\newcommand{\supervisor}{Dr Stefanos Zafeiriou}
-\newcommand{\degreetype}{Computing Science / Machine Learning}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% load some definitions and default packages
-\input{includes}
-
-% load some macros
-\input{notation}
-
-\date{September 2019}
+\title{MSc Background Report}
+\author{Peter Robertson}
+\date{}
 
 \begin{document}
-
-% load title page
-\input{titlepage}
+\maketitle
 
@@
 \pagenumbering{roman}
+\setcounter{page}{0}
 \clearpage{\pagestyle{empty}\cleardoublepage}
-\setcounter{page}{1}
-\pagestyle{fancy}
 
@@
 %--- table of contents
-\fancyhead[RE,LO]{\sffamily {Table of Contents}}
 \tableofcontents 
+\clearpage{\pagestyle{empty}\cleardoublepage}
 
-\clearpage{\pagestyle{empty}\cleardoublepage}
 \pagenumbering{arabic}
 \setcounter{page}{1}
-\fancyhead[LE,RO]{\slshape \rightmark}
-\fancyhead[LO,RE]{\slshape \leftmark}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Introduction}
+\section{Introduction}
+Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text (cite some lip reading papers).
+However there has been little progress in lip reading from 3D temporal models which contain depth information of the subject.
+A substantial barrier to this problem is the lack of 3D datasets of subjects speaking when compared to the relative abundance of video datasets.
+To first attempt this problem, new datasets which capture 3D temporal models of subjects heads speaking must be established by the community.
 
-\begin{figure}[tb]
-\centering
-\includegraphics[width = 0.4\hsize]{./figures/imperial}
-\caption{Imperial College Logo. It's nice blue, and the font is quite stylish. But you can choose a different one if you don't like it.}
-\label{fig:logo}
-\end{figure}
+\section{Lip Reading}
 
-Figure~\ref{fig:logo} is an example of a figure.
+\section{Data Generation}
+\subsection{Data Capture}
+\begin{itemize}
+    \item 2D datasets
+    \item 3D datasets
+    \item Capture New Dataset
+\end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Background}
+\subsection{Statistical Models}
+\begin{itemize}
+    \item Audio driven models
+\end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Contribution}
+\subsection{Generative Models}
+\begin{itemize}
+    \item Intro to GANs
+    \item Issues with GANs
+    \item Progression of GANs
+    \item GANs for synthesising 2D lip sync video
+\end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Experimental Results}
+\section{3D Modelling}
+\begin{itemize}
+    \item blendshapes
+    \item Model Correspondence
+\end{itemize}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Conclusion}
+\section{Recurrent Models}
+\begin{itemize}
+    \item Recurrent networks
+    \item LSTMs
+    \item Attention Mechanism
+\end{itemize}
 
-%% bibliography
-\bibliographystyle{apa}
+\section{Speech Recognition}
+Deep Speech
 
+\section{Current Work and Future Work}
+Data collection
+
+\bibliographystyle{unsrt}
+\bibliography{../ref}
 \end{document}

### 2019-05-30 Report/Progress_Report/paper_summaries.tex
tidying up repo

@@
 \subsection{Progressive Growing of GANs for Improved Quality, Stability, and Variation}
-Progressive growing GANs \cite{Karras2017b}
+Using the WGAN+GP loss function from \cite{Gulrajani2017}, Nvidia propose a novel training architecture to produce large high quality images \cite{Karras2017b}.
+Starting with a very small network which the generator and discriminator can produce and criticize 4x4 images respectively. 
+The model is trained until the discriminator is successfully fooled by the generated images.
+Then, an additional layer is 'faded' into the networks, allowing 8x8 images to processed.
+The layer is faded in with a residual connection to the previous layer, slowly weighting the new layer more heavily until the residual connection weighted to zero. 
+This avoids sudden changed in the model architecture.
+
+The two networks are mirror images of each other.
+Transpose convolution is not used, instead the image in up-sampled with 2x2 element duplication, while down-sampling is performed with average pooling.
+
+While the project at hand is not related to generating images but 3D models, the concept of working from a low dimensionality upwards, adding complexity seems intuitive. 
+In regards to the generation of faces, 4x4 and 8x8 images force the model to learn simple facial structures such as face shape which will be consistent for all samples, while higher resolution images allow the model to add finer degrees of detail to the images, but learning lower level concepts is forced initially.
+This principle isn't dissimilar to the concept of meta-learning used in \cite{Zakharov2019}.
 
 \subsection{End-to-End Speech-Driven Facial Animation with Temporal GANs}
-These guys are downstairs \cite{Vougioukas2018}
+This paper \cite{Vougioukas2018} attempts to generate video from audio, given an image of the subject.
+The generator is given an image of the subject's head and a short audio clip to generate video for.
+The model uses two discriminators, one examining individual video frames to see if they are realistic images of the subject's face, the second examines the sequence of frames produced to see if they form a realistic video sequence.
+The model produces reasonably impressive results, even more so considering that it is using the original GANs loss function from \cite{Goodfellow2014} which has been shown to be unstable, difficult to train and has been greatly improved upon though various other methods.
 
 \subsection{Self-Attention Generative Adversarial Networks}
-Attention GANs, Goodfellow involved \cite{Zhang2018}
+The issue of convolutional GANs models \cite{Radford2016} failing to produce structural pattens in an image is addressed in \cite{Zhang2018} by incorporating the attention mechanism into a GANs model.
+The paper notes that as convolutional networks require large kernels or very deep models to have a large receptive field to capture long range dependencies in an image, models can be improved with attention.
+Common examples of this issue is generating images of animals which have an incorrect number of limbs, but have realistic fur texture as this is a local dependency.
+The attention mechanism allows long range dependencies in the image to be modelled by convolutional models more easily.
+This paper is also co-authored by Ian Goodfellow, the original creator of the GANs architecture.
 
@@
 \subsection{Capture, Learning, and Synthesis of {3D} Speaking Styles}
-\cite{Cudeiro2019}
+The aim of \cite{Cudeiro2019} is to synthesise video sequences of 3D models speaking given an audio input.
+A main goal of this paper is to be independent to the speaker.
+Dataset (VOCASET) collected of 3D head scans of subjects while speaking a total of 480 sequences.
+Each sequence last 3-4 seconds for a total dataset of 29 minutes.
+Models are recorded at 60fps, resulting in 100,000 frames in the dataset.
+A common face template mesh is aligned to all the scans.
+The model uses the FLAME head model \cite{Li2017} as a template facial model.
 
-\subsection{Audio-driven facial animation by joint end-to-end learning of pose
-and emotion}
+Rather than including layers in the model to extract information on the input audio, the model uses the pre-trained Mozilla implementation of the DeepSpeech paper \cite{Hannun2014}, with a few minor improvements.
+The DeepSpeech model takes an input WAV file and returns un-normalized log probabilities for each letter and some additional characters for time slices in the audio file.
+The model outperforms all commercially available speech recognition methods at the time it was published (2014).
+
+Initially, the DeepSpeech model extracts the audio features at 50fps, this is up-sampled to 60fps.
+Then, an encoder model made up for four convolutional layers and two fully connected layers encode the audio features conditioned on the speaker so that the speaking styles of individual speakers can be learnt and later this latent space can be explored.
+Finally, a decoder made up of a fully connected layer with a linear activation function is used to output the displacements of the 5023 vertices in the FLAME face model.
+
+Personal experimentation with the completed model has shown mixed results.
+
+\subsection{Audio-driven facial animation by joint end-to-end learning of pose and emotion}
+I need to reread and take in depth notes on this.
+This paper is more impressive than I initially gave it credit for.
 Nvidia facial animation from speech \cite{Karras2017a}
 
-\subsection{End-to-end Learning for 3D Facial Animation from Raw Waveforms of
-Speech}
+\subsection{End-to-end Learning for 3D Facial Animation from Raw Waveforms of Speech}
 \cite{Pham2017}
@@
 \subsection{Synthesising 3D Facial Motion from "In-the-Wild" Speech}
+Read this again too.
 Imperial Paper \cite{Tzirakis2019}
@@
 
+\subsection{A Survey on Shape Correspondence}
+The problem of finding the correspondence between two shapes is well discussed in \cite{Kaick2011}.
+Correspondence is also referred to as registration, alignment and matching.
+The paper outlines the main differences between types of correspondence problems such as full and partial correspondence, dense and sparse correspondence.
+
+\subsection{Modelling and Correspondence of Topologically Complex 3D Shapes}
+\cite{Alhashim2015}
+
 \section{Lip Reading Papers}
@@
 \subsection{Deep Speech: Scaling up end-to-end speech recognition}
-Mozilla Deep Speech \cite{Hannun2014}
+The DeepSpeech model \cite{Hannun2014} at the time of publishing greatly outperformed all commercially available speech recognition models.
+The model uses no hand crafted components and is very robust to background noise due to the augmented training data.
+The model from the original paper was constructed and trained with efficiency in mind due to the belief that being able to train a simple but effective model for an extended period of time would yield greater results than that of a more complex model with a reduced training time.
+The model takes speech spectrograms as input and splits these into slices.
+The first three layers are fully connected, followed by a single bi-directional recurrent layer.
+The fifth layer is an additional fully connected layer, while the output layer is a softmax function.
+
+This simple model was extended by Mozilla in their publicly available implementation which made use of LSTM layer as the recurrent layer as well as using Mel-frequency cepstral coefficients (MFCC) as opposed to spectrograms.
+
+The training data was extended with data augmentation by adding noise to the audio.
+For 1000 hours of audio, 1000 hours of noise is required.
 

### 2019-06-03 Report/Progress_Report/Background.tex
Lip Reading section done

@@
 
+\usepackage{indentfirst}
 \usepackage{geometry}
@@
 \section{Introduction}
-Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text (cite some lip reading papers).
+
+\textbf{( I want to rewrite this )}
+
+Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text \cite{Chung2016}, \cite{Assael2016}, \cite{Chung2017}, \cite{Shillingford2018}.
 However there has been little progress in lip reading from 3D temporal models which contain depth information of the subject.
@@
 \section{Lip Reading}
+In recent years the problem of Lip Reading has seen huge advances due to the availability of new datasets and the use of Deep Learning models.
+In 2016 the Lip Reading Words (LRW) dataset was published \cite{Chung2016}, followed by Lip Reading Sentences (LRS) in 2017 \cite{Chung2017} and later Large-Scale Visual Speech Recognition (LSVSR) in 2018 \cite{Shillingford2018}, each of which larger in size and resulted in a model which outperforms previous models.
+Chung et al \cite{Chung2016} used a convolutional model based on the VGG-M architecture which produced character level distributions with the LRW dataset, these distributions are then processed by a language model for text prediction.
+The LRW dataset contains over 1000 hours of spoken text taken from BBC broadcasts, with a over 1000 different speakers, variations in angles and lighting, a vocabulary size of over 1000 words and over one million word utterances.
+
+The LipNet model \cite{Assael2016} was produced to be able to predict sentence-level lip reading of varied length, while previous work by \cite{Chung2016} predicted on a word level.
+The model used spatiotemporal convolutions to process multiple frames of video at once followed by a recurrent layer using Gated Recurrent Units (GRU) \cite{Cho2014}.
+The model used the GRID dataset which is of a similar size as LRW, but has far less variation in speakers.
+LipNet achieved the state of the art performance on the GRID dataset by using an model with increased complexity by incorporating 3D convolution and recurrent units.
+
+As the LRW dataset could not be used to train a model such as LipNet for sentence-level lip reading, but the GRID dataset had limitations in the number of subjects, Chung et al created the LRS dataset \cite{Chung2017}.
+The model presented was trained on both audio and video and made capable of taking either or both as the model inputs.
+To prevent the model from being dependent on a single input source, the inputs are systematically distorted or removed.
+The video input is passed through convolutional layers, followed by LSTM layers, while the audio is converted to MFCC, then input to LSTM layers.
+The two are combined with then attention mechanism and further LSTM layers and an output fully connected layer with softmax activation for character distributions.
+The model also makes use of curriculum learning by initially training the model on short sequences of single words.
+The length of training sequences are increased throughout training.
+It is stated by \cite{Chung2017} that this accelerates training and reduces overfitting.
+
+In 2018 DeepMind published their V2P model \cite{Shillingford2018} along with the LSVSR dataset which is larger than all previous datasets, containing 3886 hours of training data.
+The model follows a similar architecture to LipNet \cite{Assael2016} but with an increased number of convolutional layers and LSTM layers rather than GRUs.
+The size of the model and dataset dictated the use of 64 GPUs for training to allow a batch size of 128.
+Unlike previous models, the V2P model predicts phonemes as opposed to characters, these phonemes are then processed by a language model for word prediction as with previous models.
 
@@
 \section{Speech Recognition}
-Deep Speech
+Deep Speech \cite{Karras2017a}
 

### 2019-06-03 Report/Progress_Report/paper_summaries.tex
Lip Reading section done

@@
 \documentclass[12pt]{article}
+
+\usepackage{geometry}
+ \geometry
+ {
+     a4paper,
+     left=25mm,
+     right=25mm,
+     top=25mm,
+     bottom=25mm,
+ }
 
@@
 \section{Lip Reading Papers}
-\subsection{Out of Time: Automated Lip Sync in the Wild}
-Lip sync in the wild \cite{Chung2016}
+\subsection{Lip Reading in the Wild}
+This paper first creates a new extended dataset, later named Lip Reading Words (LRW) which allowed for a CNN architecture based on the VGG-M architecture to be trained to lip read video \cite{Chung2016}.
+The model is able to greatly outperform the state of the art models of the time.
+LRW contains over 1000 hours of spoken text with over 1000 speakers, a corpus of over 1000 words and over a million word utterances.
+
+\subsection{LipNet: Sentence-level Lipreading}
+Collaboration between Oxford and Google DeepMind \cite{Assael2016} on the GRID dataset.
+Model makes use of Spatiotemporal convolutions which convolve across multiple video frames.
+Feature maps are passed to two GRU layers and followed by a linear softmax layer for character distributions.
+CTC loss is used on output layer.
+The model is the first sentence level model, previous models have been word level.
+Able to take variable length sequences.
+
+\subsection{Lip Reading Sentences in the wild}
+\cite{Chung2017} builds on previous work to construct a model which can work on sentence level rather than just word level.
+Paper describes process of constructing a new dataset; Lip Reading Sentences (LRS).
+Model is capable of multiple inputs and is trained on both audio and video to recognise characters.
+To prevent the model from being reliant on one of the two inputs (namely the audio), the inputs are systematically distorted or removed.
+Video is passed through a CNN architecture, followed by LSTM layers.
+Audio is processed as MFCC then input to LSTM layers.
+The two are combined with then attention mechanism and further LSTM layers and an output fully connected layer with softmax activation for character distributions.
+As LSTMs are challenging to train on long sequences, curriculum learning is used.
+Initially the model is trained on short single word sequences, as training continues, the length of the input sequences increases.
+This accelerates training and reduces model overfitting.
+
+\subsection{Deep Lip Reading: a comparison of models and an online application}
+\cite{Afouras2018} adapts various models which have had success for translation and automatic speech recognition (ASR) problems.
+
+\subsection{Large-Scale Visual Speech Recognition}
+As opposed to character level predictions, \cite{Shillingford2018} predicts phonemes, combined with a language model to predict words.
+The paper describes a new dataset (LSVSR) which is larger than all previous datasets.
+Due to the size of the datasets and the model, 64 GPUs are used in parallel to train with a batch size of 64.
+The model uses five spatiotemporal convolution layers, followed by three bidirectional LSTM layers and a fully connected layer to predict phonemes.
+This model again achieves the state of the art lip reading results.
 

### 2019-06-03 Report/Progress_Report/Background.tex
mostly config

@@
-\documentclass[12pt]{article}
+\documentclass[12pt]{report}
 
+\usepackage[nottoc,notlot,notlof]{tocbibind}
 \usepackage{indentfirst}
@@
 
-\section{Introduction}
+\chapter{Introduction}
 
@@
 
-Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text \cite{Chung2016}, \cite{Assael2016}, \cite{Chung2017}, \cite{Shillingford2018}.
+Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text \cite{Chung2016, Assael2016, Chung2017, Shillingford2018}.
 However there has been little progress in lip reading from 3D temporal models which contain depth information of the subject.
@@
 
-\section{Lip Reading}
+\chapter{Lip Reading}
 In recent years the problem of Lip Reading has seen huge advances due to the availability of new datasets and the use of Deep Learning models.
@@
 
-\section{Data Generation}
-\subsection{Data Capture}
-\begin{itemize}
-    \item 2D datasets
-    \item 3D datasets
-    \item Capture New Dataset
-\end{itemize}
+\chapter{Data Generation}
+\section{Data Capture}
+Current lip reading models currently make use of 2D temporal data such as videos.
+This is convenient as video data is already widely available and relatively straightforward to process.
+The reason for the wide availability of video data is due to the ease of capturing and viewing such data.
+Unlike 2D temporal data, 3D temporal data or 3D video, requires multiple cameras to record simultaneously, which requires synchronisation, adding to the complexity of the system.
+The data must then be processed to produce the final product, whether this is in the form of video with depth information or more complex 3D scans \cite{Li2017}.
 
-\subsection{Statistical Models}
+\subsection{2D Datasets}
+Due to the large availability of video data, datasets such as LRW, LRS and LSVSR have been constructed by compiling the appropriate data.
+LRW and LRS are both comprised of content from BBC broadcasts \cite{Chung2016, Chung2017}
+
+\subsection{3D Datasets}
+
+\subsection{Original Dataset}
+
+\section{Statistical Models}
 \begin{itemize}
@@
 
-\subsection{Generative Models}
+\section{Generative Models}
 \begin{itemize}
@@
 
-\section{3D Modelling}
+\chapter{3D Modelling}
 \begin{itemize}
@@
 
-\section{Recurrent Models}
+\chapter{Recurrent Models}
 \begin{itemize}
@@
 
-\section{Speech Recognition}
+\chapter{Speech Recognition}
 Deep Speech \cite{Karras2017a}
 
-\section{Current Work and Future Work}
+\chapter{Current Work and Future Work}
 Data collection

### 2019-06-03 Report/Progress_Report/Background.tex
work on datasets

@@
-\documentclass[12pt]{report}
+\documentclass[12pt]{article}
 
 \usepackage[nottoc,notlot,notlof]{tocbibind}
+\usepackage[T1]{fontenc}
+\usepackage{hyperref}
 \usepackage{indentfirst}
@@
 
-\chapter{Introduction}
+\section{Introduction}
 
@@
 
-\chapter{Lip Reading}
+\section{Lip Reading}
 In recent years the problem of Lip Reading has seen huge advances due to the availability of new datasets and the use of Deep Learning models.
@@
 The model used spatiotemporal convolutions to process multiple frames of video at once followed by a recurrent layer using Gated Recurrent Units (GRU) \cite{Cho2014}.
-The model used the GRID dataset which is of a similar size as LRW, but has far less variation in speakers.
+The model used the GRID dataset \cite{Cooke2006} which is of a similar size as LRW, but has far less variation in speakers.
 LipNet achieved the state of the art performance on the GRID dataset by using an model with increased complexity by incorporating 3D convolution and recurrent units.
@@
 
-\chapter{Data Generation}
-\section{Data Capture}
+\section{Data Generation}
+\subsection{Data Capture}
 Current lip reading models currently make use of 2D temporal data such as videos.
@@
 
-\subsection{2D Datasets}
-Due to the large availability of video data, datasets such as LRW, LRS and LSVSR have been constructed by compiling the appropriate data.
-LRW and LRS are both comprised of content from BBC broadcasts \cite{Chung2016, Chung2017}
+\subsubsection{2D Datasets}
+Due to the large availability of video data, datasets such as GRID \cite{Cooke2006}, LRW \cite{Chung2016}, LRS \cite{Chung2017} and LSVSR \cite{Shillingford2018} have been constructed by compiling the appropriate data.
+GRID contains 34,000 samples from 34 speakers, each with 1000 sentences.
+The corpus uses sentences with a fixed grammar: 
+<command:4>, <colour:4>, <preposition:4>, <letter:25>, <number:10>, <adverb:4>, with a total vocabulary of 51 words.
 
-\subsection{3D Datasets}
+LRW and LRS are both comprised of content from BBC broadcasts \cite{Chung2016, Chung2017} allowing for a far larger corpus size of over 1000 words for LRW and over 6000 words for LRS.
+Both of datasets are captured and processed with the same pipeline summarised as follows. 
+Firstly, as the subtitles are not aligned to the video on the broadcasts, optical chapter recognition is used to obtain the text being spoken in the video clips, the audio and text are then aligned per frame using the HTK toolkit \cite{Woodland1995}.
+A HOG-based detection algorithm \cite{King2009} is used for face detection for cropping the frames to the subjects head followed by facial landmarks for mouth localisation and speaker identification. 
+\subsubsection{3D Datasets}
 
-\subsection{Original Dataset}
+\subsubsection{Original Dataset}
 
@@
 
-\chapter{3D Modelling}
+\section{3D Modelling}
 \begin{itemize}
@@
 
-\chapter{Recurrent Models}
+\section{Recurrent Models}
 \begin{itemize}
@@
 
-\chapter{Speech Recognition}
+\section{Speech Recognition}
 Deep Speech \cite{Karras2017a}
 
-\chapter{Current Work and Future Work}
+\section{Current Work and Future Work}
 Data collection

### 2019-06-04 Report/Progress_Report/Background.tex
lip reading section done.

@@
 \section{Lip Reading}
+This section shall discuss current lip reading models built with deep learning techniques and the availability of both 2D and 3D temporal datasets which lip reading models can currently be trained on.
 In recent years the problem of Lip Reading has seen huge advances due to the availability of new datasets and the use of Deep Learning models.
-In 2016 the Lip Reading Words (LRW) dataset was published \cite{Chung2016}, followed by Lip Reading Sentences (LRS) in 2017 \cite{Chung2017} and later Large-Scale Visual Speech Recognition (LSVSR) in 2018 \cite{Shillingford2018}, each of which larger in size and resulted in a model which outperforms previous models.
+Due to the large availability of video data, datasets such as GRID \cite{Cooke2006}, LRW \cite{Chung2016}, LRS \cite{Chung2017} and LSVSR \cite{Shillingford2018} have been constructed by compiling the appropriate data.
+
+Current lip reading models currently make use of 2D temporal data such as videos.
+This is convenient as video data is already widely available and relatively straightforward to obtain and process.
+Unlike 2D temporal data, 3D temporal data (3D video), requires multiple cameras to record simultaneously, which requires synchronisation, adding to the complexity of the system.
+The data must then be processed to produce the final product, whether this is in the form of video with depth information or more complex 3D scans \cite{Li2017}.
+
+\subsection{Datasets}
+
+\subsubsection{Controlled Conditions Datasets}
+The GRID dataset was released in 2006 \cite{Cooke2006} and contains 34,000 samples from 34 speakers, each with 1000 sentences.
+The corpus uses sentences with a fixed grammar: 
+<command:4>, <colour:4>, <preposition:4>, <letter:25>, <number:10>, <adverb:4>, with a total vocabulary of 51 words.
+The primary limiting factor of the GRID dataset is that all data was captured directly for the use of the dataset.
+
+\subsubsection{In The Wild Datasets}
+To build larger datasets, subsequent datasets are commonly built with data "in the wild", meaning that the videos have not been captured with the intention of being used for this dataset.
+Variations in lighting, angles, speakers and a wide vocabulary are common, this does however make the datasets more challenging to learn from.
+LRW and LRS are both comprised of content from BBC broadcasts \cite{Chung2016, Chung2017} allowing for a far larger corpus size of over 1000 words for LRW and over 6000 words for LRS.
+Both of datasets are captured and processed with the same pipeline summarised as follows. 
+Firstly, as the subtitles are not aligned to the video on the broadcasts, optical chapter recognition is used to obtain the text being spoken in the video clips, the audio and text are then aligned per frame using the HTK toolkit \cite{Woodland1995}.
+A HOG-based detection algorithm \cite{King2009} is used for face detection for cropping the frames to the subjects head followed by facial landmarks for mouth localisation and speaker identification.
+
+LSVSR is a dataset published by DeepMind and Google which makes use of the huge amount of videos on YouTube \cite{Shillingford2018}, resulting in a total length of 3886 hours of training data.
+The dataset is far larger than LRS, but aligns phonemes to frames, as opposed to words or characters.
+The pre-processing steps are similar as in LRW and LRS; alignment is performed with the algorithm laid out in previous work by DeepMind \cite{Liao2013}, faces are tracked to ensure the speaker is visible in frame.
+
+\subsection{3D Datasets}
+To the best of the author's knowledge, there are currently two datasets with 3D temporal data which are appropriate for training lip reading models.
+The first of which is LRW-3D \cite{Tzirakis2019} which has been captured from four subjects, two native English speakers and two non-native to increase variability in the dataset.
+The subjects have been captured speaking the corpus used in the LRW dataset \cite{Chung2016}, a vocabulary of 500 words.
+The resulting dataset comprises of 660 seconds of 3D meshes and audio per subject.
+The dataset is not comprised of full sentences but would be appropriate for word-level lip reading.
+
+The second dataset is the VOCASET \cite{Cudeiro2019}, captured from 6 male and 6 female subjects.
+Each subject was recorded speaking 40 sequences, each ranging from 3 to 5 seconds, resulting in a total time of 30 minutes.
+The recorded 3D meshes are registered to the FLAME model \cite{Li2017}, a statistical 3D facial mesh with around 5000 vertices.
+Unlike the LRW-3D dataset, the sequences are grammatically correct sentences, chosen to maximise phonetic diversity.
+This makes the VOCASET appropriate for creating a model for sentence-level lip reading.
+
+It should be noted that neither of these datasets were captured for the purpose of lip reading, but for synthesising realistic statistical facial models driven from an audio input, and thus the transcriptions and frames are not aligned.
+This alignment would have to be performed to the datasets before using the data to train lip reading models.
+In order to achieve this automatic speech recognition (ASR) systems such as DeepSpeech \cite{Hannun2014} could be used to solve this problem.
+
+\subsection{Lip Reading Models}
 Chung et al \cite{Chung2016} used a convolutional model based on the VGG-M architecture which produced character level distributions with the LRW dataset, these distributions are then processed by a language model for text prediction.
-The LRW dataset contains over 1000 hours of spoken text taken from BBC broadcasts, with a over 1000 different speakers, variations in angles and lighting, a vocabulary size of over 1000 words and over one million word utterances.
 
@@
 The model used spatiotemporal convolutions to process multiple frames of video at once followed by a recurrent layer using Gated Recurrent Units (GRU) \cite{Cho2014}.
-The model used the GRID dataset \cite{Cooke2006} which is of a similar size as LRW, but has far less variation in speakers.
+The model used the GRID dataset \cite{Cooke2006} in which all subjects are recorded under consistent conditions of good lighting and fixed angles.
 LipNet achieved the state of the art performance on the GRID dataset by using an model with increased complexity by incorporating 3D convolution and recurrent units.
 
-As the LRW dataset could not be used to train a model such as LipNet for sentence-level lip reading, but the GRID dataset had limitations in the number of subjects, Chung et al created the LRS dataset \cite{Chung2017}.
+As the LRW dataset could not be used to train a model such as LipNet for sentence-level lip reading, but the GRID dataset had limitations in the number of subjects and vocabulary, Chung et al created the LRS dataset \cite{Chung2017}.
 The model presented was trained on both audio and video and made capable of taking either or both as the model inputs.
@@
 The model follows a similar architecture to LipNet \cite{Assael2016} but with an increased number of convolutional layers and LSTM layers rather than GRUs.
-The size of the model and dataset dictated the use of 64 GPUs for training to allow a batch size of 128.
+It should be commented that due to the size of the model and dataset dictated the use of 64 GPUs for training to allow a batch size of 128.
 Unlike previous models, the V2P model predicts phonemes as opposed to characters, these phonemes are then processed by a language model for word prediction as with previous models.
 
+To the best of the author's knowledge, currently there do not exist any papers which have explored the use of 3D temporal datasets for the use with lip reading models. 
+This is likely due to the shortage of 3D temporal data on which such models could be trained on.
+
 \section{Data Generation}
-\subsection{Data Capture}
-Current lip reading models currently make use of 2D temporal data such as videos.
-This is convenient as video data is already widely available and relatively straightforward to process.
-The reason for the wide availability of video data is due to the ease of capturing and viewing such data.
-Unlike 2D temporal data, 3D temporal data or 3D video, requires multiple cameras to record simultaneously, which requires synchronisation, adding to the complexity of the system.
-The data must then be processed to produce the final product, whether this is in the form of video with depth information or more complex 3D scans \cite{Li2017}.
+In order to construct a deep learning lip reading model capable of being trained on 3D temporal data, appropriate datasets must be established.
+Current datasets have been captured directly \cite{Tzirakis2019, Cudeiro2019} with the use of multi-camera capture rigs under controlled situations.
+The total duration of both of these datasets is very short in comparison to the video datasets such as LRW, LRS and LSVSR, which is a limiting factor as to the models which could be trained using them.
+As the models also use different mesh models to represent the data that has been captured this also prevents the two datasets being joined directly.
+Unlike video data, there currently lacks a large body of 3D video data which is publicly available, limiting the construction of 3D datasets to directly capturing more 3D scans with multi-camera capture rigs, similar to those used in \cite{Tzirakis2019, Cudeiro2019} and generating synthetic training data.
 
-\subsubsection{2D Datasets}
-Due to the large availability of video data, datasets such as GRID \cite{Cooke2006}, LRW \cite{Chung2016}, LRS \cite{Chung2017} and LSVSR \cite{Shillingford2018} have been constructed by compiling the appropriate data.
-GRID contains 34,000 samples from 34 speakers, each with 1000 sentences.
-The corpus uses sentences with a fixed grammar: 
-<command:4>, <colour:4>, <preposition:4>, <letter:25>, <number:10>, <adverb:4>, with a total vocabulary of 51 words.
+\subsection{Audio Driven Data Generation}
 
-LRW and LRS are both comprised of content from BBC broadcasts \cite{Chung2016, Chung2017} allowing for a far larger corpus size of over 1000 words for LRW and over 6000 words for LRS.
-Both of datasets are captured and processed with the same pipeline summarised as follows. 
-Firstly, as the subtitles are not aligned to the video on the broadcasts, optical chapter recognition is used to obtain the text being spoken in the video clips, the audio and text are then aligned per frame using the HTK toolkit \cite{Woodland1995}.
-A HOG-based detection algorithm \cite{King2009} is used for face detection for cropping the frames to the subjects head followed by facial landmarks for mouth localisation and speaker identification. 
-\subsubsection{3D Datasets}
-
-\subsubsection{Original Dataset}
-
-\section{Statistical Models}
+\subsection{Adversarial Trained Generative Methods}
 \begin{itemize}
-    \item Audio driven models
-\end{itemize}
-
-\section{Generative Models}
-\begin{itemize}
-    \item Intro to GANs
     \item Issues with GANs

### 2019-06-04 Report/Progress_Report/Background.tex
voca paper done, starting GANS

@@
 \usepackage{indentfirst}
+\usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{geometry}
@@
 
-\subsection{3D Datasets}
+\subsection{3D Datasets} \label{3D Datasets}
 To the best of the author's knowledge, there are currently two datasets with 3D temporal data which are appropriate for training lip reading models.
@@
 
+\subsubsection{Imperial}
+\cite{Tzirakis2019}
+
+\subsubsection{VOCA}
+The VOCA model \cite{Cudeiro2019} synthesises video sequences of 3D models speaking given an audio input.
+The VOCASET dataset discussed in section \ref{3D Datasets}, was captured with the intention of training this model to be independent of the speaker, hence a large range of speakers are used within the dataset.
+The model is comprised of three sections: audio feature extraction, a feature encoder and a decoder to drive a template facial mesh from the FLAME model \cite{Li2017}.
+The audio feature extraction makes use of the pre-trained Mozilla implementation of the DeepSpeech model, based on the paper by Hannun et al. \cite{Hannun2014}.
+The DeepSpeech model takes audio as an input and returns the unnormalised log-probabilities for an alphabet of the 26 standard characters, a space, apostrophe and blank character for time slices in the audio input.
+The encoder is a convolutional network which is conditioned on the speakers identity, such that the latent space of speaker styles can later be explored on new audio inputs.
+Finally, the decoder is made up of a fully connected layer with a linear activation function is used to output the displacements of the 5023 vertices in the template face.
+
+\subsubsection{Nvidia}
+\cite{Karras2017a}
+
 \subsection{Adversarial Trained Generative Methods}
+A recent development in generating synthetic data samples is the use of generative adversarial networks \cite{Goodfellow2014}.
+The concept behind generative adversarial networks is two have two machine learning models; a generator and a discriminator.
+The task of the generator is to produce samples from an unknown high order probability distribution which correctly resemble samples from the distribution defined by training data.
+The generator achieves this by transforming a random sample from a known probability distribution, such as a Gaussian distribution as used in \cite{Goodfellow2014}, to a sample from the unknown distribution.
+This is achieved by finding the function which maps between the two distributions.
+The discriminator however, attempts to correctly learn to discriminate between the real and the fake generated samples.
+
+\begin{equation} \label{cgans}
+    \mathcal{L}_{adv} = \mathbb{E}_{x \sim X_{real}}[\log(D_{src}(x)] 
+    + \mathbb{E}_{x \sim X_{real},c \sim C}[\log(1-D_{src}(G(x,c))]
+\end{equation}
+
+This paper uses two feed forward neural networks (as proof of concept) as generator and discriminator.
+The original loss function is used, although it is mentioned that even at this stage issues with it are highlighted.
+Argument made to say that while learning a probability distribution is great, the goal is to produce samples from it, so generating sample from an implicit distribution is as useful.
+No need to go into extreme depth as this was covered in the Deep Learning lectures.
 \begin{itemize}

### 2019-06-04 Scripts/equation_plotter.py
handy scripts

@@ -1,0 +1,27 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import os
+
+# example data
+x = np.linspace(0.01, 0.99, 100)
+
+# Equations to plot
+y1 = np.log(x)
+y2 = np.log(1-x)
+
+fig, ax = plt.subplots()
+ax.set_title('Loss Plots for Original GANs Loss Function')
+
+line1, = ax.plot(x, y1, label='Discriminator Loss')
+line2, = ax.plot(x, y2, label='Generator Loss')
+
+ax.legend()
+#plt.show()
+
+filename = "original_gans_loss.png"
+save_path = os.path.join("figures", filename)
+
+if not os.path.exists("figures"):
+    os.makedirs("figures")
+
+plt.savefig(save_path)

### 2019-06-04 Report/Progress_Report/Background.tex
GANs gen loss plots

@@
 \usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{subcaption}
 \usepackage{geometry}
@@
      right=25mm,
-     top=25mm,
-     bottom=25mm,
+     top=30mm,
+     bottom=30mm,
  }
@@
 A recent development in generating synthetic data samples is the use of generative adversarial networks \cite{Goodfellow2014}.
-The concept behind generative adversarial networks is two have two machine learning models; a generator and a discriminator.
+The concept behind generative adversarial networks (GANs) is two have two machine learning models; a generator and a discriminator.
 The task of the generator is to produce samples from an unknown high order probability distribution which correctly resemble samples from the distribution defined by training data.
@@
 
-\begin{equation} \label{cgans}
-    \mathcal{L}_{adv} = \mathbb{E}_{x \sim X_{real}}[\log(D_{src}(x)] 
-    + \mathbb{E}_{x \sim X_{real},c \sim C}[\log(1-D_{src}(G(x,c))]
+\begin{equation} \label{gans_loss}
+    \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
+                              + \mathbb{E}_{z \sim p_{z}(z)} [\log (1 - D(G(x)))]
 \end{equation}
 
-This paper uses two feed forward neural networks (as proof of concept) as generator and discriminator.
-The original loss function is used, although it is mentioned that even at this stage issues with it are highlighted.
-Argument made to say that while learning a probability distribution is great, the goal is to produce samples from it, so generating sample from an implicit distribution is as useful.
-No need to go into extreme depth as this was covered in the Deep Learning lectures.
+\noindent The original loss function (\ref{gans_loss}) proposed by Ian Goodfellow forms a min-max game, where the loss of the generator is attempting to be minimised by having the discriminator label all the generated samples as real.
+While the loss of the discriminator is maximised by correctly classifying real and fake samples.
+The two networks are trained in an alternating fashion until the discriminator achieves an accuracy of 50\%, effectively making binary guesses between real and generated samples.
+
+GANs however, are difficult to train for two main reasons.
+Firstly, the equation (\ref{gans_loss}) is challenging as it provides small gradients while generated samples are poor as discussed in \cite{Goodfellow2014}, making training difficult.
+Progress in developing new loss functions is discussed in section \ref{Stability_to_GANs}
+Secondly, early networks must also be balanced with a similar model capacity to prevent one from getting too much better than the other, preventing the other from improving.
+Various architectural changes have improved this issue \cite{Radford2016, Zhang2018}, although it seems to be closely tied to the loss function being used \cite{Gulrajani2017}.
+
+\subsubsection{Stability Improvements of GANs Networks} \label{Stability_to_GANs}
+A primary research focus around GANs has been in finding new loss functions on which to train the model to improve stability and performance.
+The original loss function proposed in the original paper \cite{Goodfellow2014} identifies an issue with equation (\ref{gans_loss}) in that the gradient back propagated to the generator when the generated samples are poor, is very low as shown in figure \ref{fig:Goodfellow_plot}.
+
+\begin{figure}[h]
+    \centering
+        \includegraphics[width=0.8\textwidth]{../figures/goodfellow_gen_losses.png}
+    \caption{Goodfellow's Generator Loss Plots \cite{Goodfellow2014}}\label{fig:Goodfellow_plot}
+\end{figure}
+
+\noindent This in turn makes it very challenging to improve the performance of the generator.
+The suggested improvement made in \cite{Goodfellow2014} is rather than to maximise the the number of generated samples which the discriminator incorrectly classifies, to minimise the number of generator samples the discriminator correctly classifies, described in equation (\ref{gans_loss2}).
+Figure \ref{fig:Goodfellow_plot} shows how this results in the gradient propagated back to the generator is far larger when the generated samples are poor.
+
+\begin{equation} \label{gans_loss2}
+    \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
+                              - \mathbb{E}_{z \sim p_{z}(z)} [\log (D(G(x)))]
+\end{equation}
+
 \begin{itemize}
-    \item Issues with GANs
-    \item Progression of GANs
-    \item GANs for synthesising 2D lip sync video
+    \item Batch Norm
+    \item WGAN
+    \item WGAN+GP
+\end{itemize}
+
+\subsubsection{Architectural Developments}
+\begin{itemize}
+    \item Conditional GANs
+    \item DCGAN
+    \item Temporal GANs
+    \item Attention GANs
 \end{itemize}

### 2019-06-04 Scripts/equation_plotter.py

@@ -6,20 +6,21 @@ import os
 x = np.linspace(0.01, 0.99, 100)
 
 # Equations to plot
-y1 = np.log(x)
-y2 = np.log(1-x)
+y1 = np.log(1-x)
+y2 = -np.log(x)
 
 fig, ax = plt.subplots()
-ax.set_title('Loss Plots for Original GANs Loss Function')
+#ax.set_title('Loss Plots for Original GANs Loss Function')
 
-line1, = ax.plot(x, y1, label='Discriminator Loss')
-line2, = ax.plot(x, y2, label='Generator Loss')
+line1, = ax.plot(x, y1, label='Original Generator Loss')
+line2, = ax.plot(x, y2, label='Improved Generator Loss')
 
 ax.legend()
 #plt.show()
 
-filename = "original_gans_loss.png"
-save_path = os.path.join("figures", filename)
+filename = "goodfellow_gen_losses.png"
+dir_path = os.path.dirname(os.path.realpath(__file__))
+save_path = os.path.join(dir_path, "figures", filename)
 
 if not os.path.exists("figures"):
     os.makedirs("figures")

### 2019-06-05 Report/Progress_Report/Background.tex
progress on WGAN section

@@
 
-\begin{equation} \label{gans_loss}
+\begin{equation} \label{eq:gans_loss}
     \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
@@
 
-\noindent The original loss function (\ref{gans_loss}) proposed by Ian Goodfellow forms a min-max game, where the loss of the generator is attempting to be minimised by having the discriminator label all the generated samples as real.
+\noindent The original loss function (\ref{eq:gans_loss}) proposed by Ian Goodfellow forms a min-max game, where the loss of the generator is attempting to be minimised by having the discriminator label all the generated samples as real.
 While the loss of the discriminator is maximised by correctly classifying real and fake samples.
@@
 GANs however, are difficult to train for two main reasons.
-Firstly, the equation (\ref{gans_loss}) is challenging as it provides small gradients while generated samples are poor as discussed in \cite{Goodfellow2014}, making training difficult.
+Firstly, the equation (\ref{eq:gans_loss}) is challenging as it provides small gradients while generated samples are poor as discussed in \cite{Goodfellow2014}, making training difficult.
 Progress in developing new loss functions is discussed in section \ref{Stability_to_GANs}
@@
 \subsubsection{Stability Improvements of GANs Networks} \label{Stability_to_GANs}
+There have been a large number of papers presenting new techniques with differing levels of success and training stability, a small handful of key papers which provided large advances in the generative adversarial training model shall be discussed here.
 A primary research focus around GANs has been in finding new loss functions on which to train the model to improve stability and performance.
-The original loss function proposed in the original paper \cite{Goodfellow2014} identifies an issue with equation (\ref{gans_loss}) in that the gradient back propagated to the generator when the generated samples are poor, is very low as shown in figure \ref{fig:Goodfellow_plot}.
+The original loss function proposed in the original paper \cite{Goodfellow2014} identifies an issue with equation (\ref{eq:gans_loss}) in that the gradient back propagated to the generator when the generated samples are poor, is very low as shown in figure \ref{fig:Goodfellow_plot}.
 
@@
 \noindent This in turn makes it very challenging to improve the performance of the generator.
-The suggested improvement made in \cite{Goodfellow2014} is rather than to maximise the the number of generated samples which the discriminator incorrectly classifies, to minimise the number of generator samples the discriminator correctly classifies, described in equation (\ref{gans_loss2}).
+The suggested improvement made in \cite{Goodfellow2014} is rather than to maximise the the number of generated samples which the discriminator incorrectly classifies, but to minimise the number of generator samples the discriminator correctly classifies, described in equation (\ref{eq:gans_loss2}).
 Figure \ref{fig:Goodfellow_plot} shows how this results in the gradient propagated back to the generator is far larger when the generated samples are poor.
 
-\begin{equation} \label{gans_loss2}
+\begin{equation} \label{eq:gans_loss2}
     \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
@@
 
-\begin{itemize}
-    \item Batch Norm
-    \item WGAN
-    \item WGAN+GP
-\end{itemize}
+Further stability improvements were proposed in \cite{Radford2016} which allowed deep convolutional generative adversarial networks (DCGANs) to be successfully trained for the first time.
+Radford et al. proposed three main contributions.
+Replace deterministic pooling layers with strided convolutions in both the generator and discriminator networks, allowing the networks to learn their own spatial upsampling and downsampling.
+Remove all fully connected layers used on top of convolutional layers, resulting in a fully convolutional model.
+Apply batch normalisation \cite{Ioffe2015} before the input of each layer.
+This normalises the input to each unit to zero mean and unit variance.
+This assists with training problems due to poor weight initialisation and allows gradients to flow through deeper networks more easily.
+Radford et al. state this to be a critical improvement to allow generator networks to begin learning by preventing all samples from collapsing to a single point.
+
+In an attempt to stabilize the training of GANs models, the Wasserstein or 'Earth-Mover' loss function was proposed by Arjovsky et al. \cite{Arjovsky2017}.
+By using the Wasserstein loss function \ref{eq:wgan}, the generator and discriminator no longer have to alternate training epochs to try to maintain model stability, but can both be trained to saturation in a loop.
+
+\begin{equation} \label{eq:wgan}
+    \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
+                              + \mathbb{E}_{z \sim p_{z}(z)} [\log (1 - D(G(x)))]
+\end{equation}
+
+\noindent By using the Wasserstein loss function, the discriminator, or critic as it is named by Arjovsky et al., is trained to an optimum point, at which time the loss provided to the generator is at a maximum, making training the generator easier.
+The discriminator is referred to as the critic as it is no longer attempting to classify samples are real or fake.
+The generator can then be trained to an optimum position, at which time, training the critic can continue again.
+This method has greatly improved stability and is fairly robust to a large range of model architectures which have been tested.
+The generator and critic also no longer have to have a similar capacity.
+
+To achieve this, the Lipschtiz condition must be met.
+This is achieved by clipping the weights on the model.
+The paper points out that this method of achieving the Lipschtiz condition is non-ideal and is a point of further work.
+
+Other points of note, the loss plot produced by the WGAN model seems to correlate with perceived image quality, which could not be said for precious loss functions.
+
+\subsubsection{WGAN}
+\subsubsection{WGAN+GP}
 

### 2019-06-05 Report/Progress_Report/Background.tex
main gans section complete.

@@
 \documentclass[12pt]{article}
+
+% some macros
+\input{../notation}
 
@@
 \begin{equation} \label{eq:gans_loss}
-    \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
-                              + \mathbb{E}_{z \sim p_{z}(z)} [\log (1 - D(G(x)))]
+    \min_{G} \max_{D} V(G, D) = \E{\bm{x} \sim p_{data}(\bm{x})} [\log D(\bm{x})]
+                              + \E{\bm{z} \sim p_{z}(\bm{z})} [\log (1 - D(G(\bm{z})))]
 \end{equation}
@@
 \begin{equation} \label{eq:gans_loss2}
-    \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
-                              - \mathbb{E}_{z \sim p_{z}(z)} [\log (D(G(x)))]
+    \min_{G} \max_{D} V(G, D) = \E{\bm{x} \sim p_{data}(\bm{x})} [\log D(\bm{x})]
+                              - \E{\bm{z} \sim p_{z}(\bm{z})} [\log (D(G(\bm{z})))]
 \end{equation}
@@
 
-In an attempt to stabilize the training of GANs models, the Wasserstein or 'Earth-Mover' loss function was proposed by Arjovsky et al. \cite{Arjovsky2017}.
-By using the Wasserstein loss function \ref{eq:wgan}, the generator and discriminator no longer have to alternate training epochs to try to maintain model stability, but can both be trained to saturation in a loop.
+In an attempt to stabilize the training of GANs models, the Wasserstein or 'Earth Mover Distance' loss function was proposed by Arjovsky et al.
+\cite{Arjovsky2017} shown in equation (\ref{eq:wgan}) where $\mathcal{D}$ is a set of 1-Lipschtiz functions.
 
 \begin{equation} \label{eq:wgan}
-    \min_{G} \max_{D} V(G, D) = \mathbb{E}_{x \sim p_{data}(x)} [\log D(x)]
-                              + \mathbb{E}_{z \sim p_{z}(z)} [\log (1 - D(G(x)))]
+    \min_{G} \max_{D \in \mathcal{D}} W(\Pd{r}, \Pd{g}) =
+            \E{\bm{x} \sim \Pd{r}} [D(\bm{x})]
+            -\E{\bm{\hat{x}} \sim \Pd{g}} [D(\hat{\bm{x}})]
 \end{equation}
 
-\noindent By using the Wasserstein loss function, the discriminator, or critic as it is named by Arjovsky et al., is trained to an optimum point, at which time the loss provided to the generator is at a maximum, making training the generator easier.
-The discriminator is referred to as the critic as it is no longer attempting to classify samples are real or fake.
-The generator can then be trained to an optimum position, at which time, training the critic can continue again.
-This method has greatly improved stability and is fairly robust to a large range of model architectures which have been tested.
-The generator and critic also no longer have to have a similar capacity.
+\noindent The WGAN model uses a critic as opposed to a discriminator, this is due to the fact that the discriminator is no longer a binary classifier, but being used to critique the real and generated samples.
+As the Wasserstein function is continuous and differentiable, the critic can be trained until optimality, and \cite{Arjovsky2017} argues that it should be.
+As the critic is trained to optimality it does not saturate, but converges to a linear function.
+This provides the generator with a gradient which is more reliable, resulting in the generator learning consistently.
+The Wasserstein loss function has been shown to greatly improve training stability by providing consistent gradients throughout training and no longer requiring that the two networks have a balanced model capacity.
 
-To achieve this, the Lipschtiz condition must be met.
-This is achieved by clipping the weights on the model.
-The paper points out that this method of achieving the Lipschtiz condition is non-ideal and is a point of further work.
+In order to use the Wasserstein distance as the loss for WGAN, the Lipschtiz condition must be enforced.
+In the WGAN the model weights are clipped to enforce this constraint \cite{Arjovsky2017}, however this is stated to be a non-ideal method of achieving this constraint and calls for future work to be investigate more effective methods.
+The use of gradient penalty is proposed in \cite{Gulrajani2017} in order to satisfy this condition more elegantly as described in equation (\ref{eq:wgan_gp}).
 
-Other points of note, the loss plot produced by the WGAN model seems to correlate with perceived image quality, which could not be said for precious loss functions.
+\begin{equation} \label{eq:wgan_gp}
+    \min_{G} \max_{D \in \mathcal{D}} W(\Pd{r}, \Pd{g}) 
+        = \E{\bm{x} \sim \Pd{r}} [D(\bm{x})]
+        - \E{\bm{\tilde{x}} \sim \Pd{g}} [D(\tilde{{\bm{x}}})]
+        + \lambda \E{\bm{\hat{x}} \sim \Pd{\bm{\hat{x}}}} 
+            [(\| \nabla_{\bm{\hat{x}}} D(\bm{\hat{x}}) \| - 1)^2]
+\end{equation}
 
-\subsubsection{WGAN}
-\subsubsection{WGAN+GP}
+\noindent Using the Wasserstein loss function with gradient penalty enforces the Lipschtiz condition and allows highly complex architectures to be trained successfully, including those with residual units \cite{Gulrajani2017}.
 

### 2019-06-06 Report/Progress_Report/Background.tex
finished generative section

@@
 \subsubsection{Architectural Developments}
-\begin{itemize}
-    \item Conditional GANs
-    \item DCGAN
-    \item Temporal GANs
-    \item Attention GANs
-\end{itemize}
+Mirza et al. showed that GANs could be conditioned on additional inputs to the generator network as the noise component \cite{Mirza2014}.
+This has allowed facilitated other uses of GANs such as image to image translation for style transfer \cite{Zhu2017}.
+Vougioukas et al.
+used a temporal model to generate a sequence of video frames given an image of a subject and an audio sequence of spoken text to synthesise the subject speaking \cite{Vougioukas2018}.
+The model uses two discriminators, one to determine if individual frames are realistic images of the subject's face and a second which evaluates the sequence of video frames to determine if it is realistic.
+The model uses temporal components to examine if the frames of video are consistent in time, preventing sudden jumps in facial position.
+
+Other novel architectures include the progressively growing GAN model \cite{Karras2017b} which is able to generate highly realistic images of faces to a high resolution.
+This is achieved by initially training a shallow model to produce 4x4 pixel images before increasing the depth of the network and training further at a higher resolution.
+By forcing the model to firstly produce and examine low resolution images the model firstly has to be able to synthesise simple low level features effectively, such as facial shape which are common to all samples.
+Once the model can produce these features a higher resolution is used, allowing it to learn more complex features, such as hair and eyes.
+
+The attention mechanism \cite{Vaswani2017} has been shown to be usefully when applied to image data \cite{Xu2015} in order to capture relationships between spatially distant points.
+As such points are further apart in the image, previously deeper convolutional models were required to allow a large enough receptive field to capture information on these two points.
+Zhang et al.
+point out that this is a common issue with DCGANs \cite{Radford2016}, where generated samples fail to produce structural patterns, while they do exceedingly well at local textural patterns.
+This is often seen in generated images of animals with realistic fur, but oddly shaped or an incorrect number of limbs.
+The attention mechanism is applied to GANs \cite{Zhang2018} to allow for long range dependencies in the image to be modelled by convolutional models more effectively.
 

### 2019-06-06 Report/Progress_Report/Background.tex
Finished section on data generation

@@
 \setcounter{page}{1}
+\frenchspacing
 
@@
 
+\section{3D Modelling}
+\begin{itemize}
+    \item blendshapes
+    \item Model Correspondence
+\end{itemize}
+
 \section{Data Generation}
@@
 \subsection{Audio Driven Data Generation}
+Karras et al proposed a method for generating 3D facial animation from audio with the use of a CNN architecture \cite{Karras2017a}.
+The model is actor specific, but is only trained on 3-5 minutes of data of the actor.
+Short range temporal features are first extracted from the audio by a formant analysis network.
+This representation is then analysed by an articulation network which also accepts a learned emotional state of the speaker.
+The output layer drives displacements from a neutral 3d mesh of the actor.
 
-\subsubsection{Imperial}
-\cite{Tzirakis2019}
+However, as the model by Karras et al. is not independent of the actor it cannot generalise to new subjects.
+Tzirakis et al. propose a model which is independent of speaker and capture rig \cite{Tzirakis2019}.
+As discussed in section \ref{3D Datasets}, a dataset was constructed of 3D speaking faces using the LRW dataset \cite{Chung2016} for the corpus.
+This allowed Tzirakis et al. to create a model which can synthesise facial motion from audio from the LRW dataset.
+The model used is similar to that used in \cite{Karras2017a}, firstly extracting short term temporal features from the input audio with a convolutional network followed by another convolutional network to analyse the extracted features.
+Unlike the model used in \cite{Karras2017a}, the model used in \cite{Tzirakis2019} is trained to drive learnt blendshapes.
+This reduces the number of output parameters of the model substantially in comparison to that used by Karras et al.
 
-\subsubsection{VOCA}
-The VOCA model \cite{Cudeiro2019} synthesises video sequences of 3D models speaking given an audio input.
+Similar to \cite{Tzirakis2019}, the VOCA model \cite{Cudeiro2019} synthesises video sequences of 3D models speaking given an audio input.
 The VOCASET dataset discussed in section \ref{3D Datasets}, was captured with the intention of training this model to be independent of the speaker, hence a large range of speakers are used within the dataset.
@@
 Finally, the decoder is made up of a fully connected layer with a linear activation function is used to output the displacements of the 5023 vertices in the template face.
-
-\subsubsection{Nvidia}
-\cite{Karras2017a}
 
@@
 
-\section{3D Modelling}
-\begin{itemize}
-    \item blendshapes
-    \item Model Correspondence
-\end{itemize}
-
-\section{Recurrent Models}
-\begin{itemize}
-    \item Recurrent networks
-    \item LSTMs
-    \item Attention Mechanism
-\end{itemize}
-
-\section{Speech Recognition}
-Deep Speech \cite{Karras2017a}
+To the best of the author's knowledge, there currently exists no generative adversarial networks which aim to generate 3D facial models with speech audio as a conditional input.
 

### 2019-06-06 Report/Progress_Report/paper_summaries.tex
Finished section on data generation

@@
 \subsection{Audio-driven facial animation by joint end-to-end learning of pose and emotion}
-I need to reread and take in depth notes on this.
-This paper is more impressive than I initially gave it credit for.
 Nvidia facial animation from speech \cite{Karras2017a}
+Attempts to generate 3D facial animation from audio.
+Output vertices difference from neutral face mesh.
+Specific to actor, trained on 3-5 minutes of footage for that actor.
+Model extracts temporal audio features with convolutional net.
+Then convolutional net to analyse temporal components of speech.
+Articulation network also accepts emotional state input to aim ambiguities between different facial shape from same audio.
 
@@
 \subsection{Synthesising 3D Facial Motion from "In-the-Wild" Speech}
-Read this again too.
+Model is independent of speaker and capture rig.
+Create dataset by recording talking 4D faces speaking words from LRW dataset.
+Register 3D meshes and learn 3D blendshapes.
+Use DCAW to align speech with meshes to give ground truth.
+Train a model to predict 3D facial motion from speech under uncontrolled conditions. 
+Use audio from LRW dataset to train with.
 Imperial Paper \cite{Tzirakis2019}

### 2019-06-06 Report/Progress_Report/Background.tex
completed background report

@@
 \input{../notation}
+\input{includes}
 
-\usepackage[nottoc,notlot,notlof]{tocbibind}
-\usepackage[T1]{fontenc}
-\usepackage{hyperref}
-\usepackage{indentfirst}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{geometry}
- \geometry
- {
-     a4paper,
-     left=25mm,
-     right=25mm,
-     top=30mm,
-     bottom=30mm,
- }
+%\title{MSc Background Report}
+%\author{Peter Robertson}
+%\date{}
 
-\title{MSc Background Report}
-\author{Peter Robertson}
-\date{}
+\newcommand{\reporttitle}{Background and Progress Report}
+\newcommand{\reportauthor}{Peter Robertson}
+\newcommand{\supervisor}{Dr Stefanos Zafeiriou}
+\newcommand{\degreetype}{Computing Science / Machine Learning}
 
 \begin{document}
-\maketitle
+%\maketitle
+\input{titlepage}
 
@@
 
-\textbf{( I want to rewrite this )}
-
-Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text \cite{Chung2016, Assael2016, Chung2017, Shillingford2018}.
-However there has been little progress in lip reading from 3D temporal models which contain depth information of the subject.
-A substantial barrier to this problem is the lack of 3D datasets of subjects speaking when compared to the relative abundance of video datasets.
+Visual speech prediction, often referred to as lip reading, is a very difficult task, partly due to ambiguities between phonemes such as '\textit{p}' and '\textit{b}', which look the same but sound different.
+Current methods of addressing this problem focus on training deep learning models using 2D temporal data of people speaking with the relevant video frames labelled with the correct spoken text \cite{Chung2016, Assael2016, Chung2017, Shillingford2018}.
+There currently do not exist any models which attempt to address this problem with the use of 3D temporal datasets made up of head scans of subjects, capturing depth and thus, more information of the mouth which may be of use to the models.
+However, there currently exists a severe lack of availability of such datasets due to the complexity and hardware requirements in capturing such data.
+Currently their exists two publicly available datasets: LRW-3D \cite{Tzirakis2019} and the VOCASET \cite{Cudeiro2019}, both of which are relatively small in comparison to the models used in current lip reading models.
 To first attempt this problem, new datasets which capture 3D temporal models of subjects heads speaking must be established by the community.
+As well as capturing new data, expanding existing datasets with generated synthetic data with the use of generative adversarial networks may assist in the solution to this problem.
 
@@
 
-\section{3D Modelling}
-\begin{itemize}
-    \item blendshapes
-    \item Model Correspondence
-\end{itemize}
+\section{3D Facial Modelling}
+Unlike 2D video where the subject is captured from a single angle without any depth information, 3D models cannot be captured with a single camera.
+In most instances, 3D head scans are captured with the use of a multi-camera rig with multiple cameras capturing video simultaneously.
+The number and type of camera can vary from system to system, but the principle remains consistent.
+All cameras must be synchronised to capture footage at the same time, with the same frame rate.
+This is inherently a more complex system than 2D video capture as is therefore more expensive, limiting the accessibility of such systems.  
+Once the images have been captured, the footage must be processed in order to produce a facial mesh of the subject \cite{Li2017}.
+
+\subsection{Blendshapes}
+A facial mesh is made up of a point cloud of vertices and vertex connections.
+Depending on the mesh capture pipeline, the number of vertices may vary, but even a low resolution mesh may contain several thousand vertices.
+Capturing realistic facial motion by modelling the displacements of all of these vertices is impractical when manipulating a model by hand.
+Blendshapes attempt to model aspects of realistic facial motion by finding the relations between these vertices and manipulating them simultaneously \cite{Lewis2010}, by reducing the number of parameters, the model becomes more manageable to control.
+One method of producing model blendshapes is with principle component analysis (PCA).
+By applying PCA on a set of facial meshes, the resulting principle components represent the changes in facial motion which capture the largest amount of variation in the set while reducing the number of model parameters substantial.
 
@@
 \end{equation}
+\quad
 
@@
 \end{figure}
+\quad
 
@@
 \section{Current Work and Future Work}
-Data collection
+At the current stage of the project, the current goal is to construct a new dataset of 3D facial scans of subjects speaking selected phrases by collecting the relevant data, this dataset can be combined with LRW-3D and VOCASET.
+In order to achieve this, the facial meshes from the individual datasets must be brought into correspondence, creating a larger dataset.
+The current issue with models which are to be trained on 3D temporal facial scans is the lack of such datasets due to the complexity in creating such datasets.
+By combining new and existing datasets, this aids in the solution of this issue.
+
+Future work will include training a GANs model which is capable to driving a template 3D mesh given an conditional audio input.
+This will allow for synthetic data augmentation to be applied to the current 3D facial datasets.
+This augmented dataset can be used to train further models which accept 3D facial meshes as input for tasks such as lip reading in order to see if model performance can be enhanced without the need for additional data capture.
+
+\newpage
 

### 2019-06-08 src/mesh/hello_mesh.py
hacky methods for extracting mesh data from ply files with plyfile

@@ -1,0 +1,43 @@
+from plyfile import PlyData
+import os
+import numpy as np
+import time
+
+filename = "head.ply"
+dir_path = os.path.dirname(os.path.realpath(__file__))
+file_path = os.path.join(dir_path, filename)
+ply_files = "/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01"
+
+def ply_verts_to_array(ply_verts):
+    length = ply_verts.count
+    verts = np.empty((length, 3))
+
+    for vert in range(length):
+        v = ply_verts[vert]
+        (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
+        verts[vert] = [x, y, z]
+
+    return verts
+
+def list_of_plys(filepath):
+    """
+    Returns a list of ply files in given directory
+    """
+    files = []
+    for file in os.listdir(filepath):
+        if file.endswith("1.ply"):
+            files.append(file)
+    return files
+
+verts = []
+
+files = list_of_plys(ply_files)
+for file in files:
+    filepath = os.path.join(ply_files, file)
+
+    with open(filepath, 'rb') as f:
+        plydata = PlyData.read(f)
+        face_verts = ply_verts_to_array(plydata['vertex'])
+    verts.append(face_verts)
+
+print(len(verts))

### 2019-06-10 src/mesh/pca_mesh.py
pca mesh verticies to create blendshapes

@@ -1,0 +1,63 @@
+from plyfile import PlyData
+from sklearn.decomposition import PCA
+import os
+import numpy as np
+import time
+
+class Mesh_Collect():
+    def __init__(self, ply_files):
+        self.ply_files = ply_files
+        self._num_vertices = 5023
+        self.num_files = 0
+
+    def ply_verts_to_array(self, ply_verts):
+        length = ply_verts.count
+        verts = np.empty((length, 3))
+
+        for vert in range(length):
+            v = ply_verts[vert]
+            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
+            verts[vert] = [x, y, z]
+
+        return verts
+
+    def list_of_plys(self):
+        """
+        Returns a list of ply files in given directory
+        """
+        files = []
+        for file in os.listdir(self.ply_files):
+            if file.endswith(".ply"):
+                files.append(file)
+        
+        self.num_files = len(files)
+        return files
+
+    def collect_ply_data(self, ply_verts, target_array, array_postion):
+        for vert in range(self._num_vertices):
+            v = ply_verts[vert]
+            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
+            target_array[array_postion+vert] = [x, y, z]
+
+ply_files = "/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01"
+
+ms = Mesh_Collect(ply_files)
+
+files = ms.list_of_plys()
+
+verts = np.zeros((ms._num_vertices * ms.num_files, 3))
+position = 0
+for file in files:
+    filepath = os.path.join(ply_files, file)
+
+    with open(filepath, 'rb') as f:
+        plydata = PlyData.read(f)
+        ms.collect_ply_data(plydata['vertex'], verts, position)
+    position += ms._num_vertices
+
+pca = PCA(n_components=3)
+pca.fit(verts)  
+
+print(pca.explained_variance_ratio_)  
+print(pca.singular_values_) 
+print(pca.components_)

### 2019-06-10 src/mesh/pca_mesh.py
pca blendshapes script finished

@@ -1,63 +1,72 @@
 from plyfile import PlyData
-from sklearn.decomposition import PCA
+import scipy.linalg as la
 import os
 import numpy as np
 import time
 
-class Mesh_Collect():
+class Blendshapes():
     def __init__(self, ply_files):
         self.ply_files = ply_files
         self._num_vertices = 5023
         self.num_files = 0
 
-    def ply_verts_to_array(self, ply_verts):
-        length = ply_verts.count
-        verts = np.empty((length, 3))
-
-        for vert in range(length):
-            v = ply_verts[vert]
-            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
-            verts[vert] = [x, y, z]
-
-        return verts
-
     def list_of_plys(self):
         """
         Returns a list of ply files in given directory
         """
-        files = []
-        for file in os.listdir(self.ply_files):
-            if file.endswith(".ply"):
-                files.append(file)
+        selected_files = []
+        for root, _dirs, files in os.walk(self.ply_files):
+            for file in files:
+                if file.endswith("1.ply"):
+                    selected_files.append(os.path.join(root, file))
         
-        self.num_files = len(files)
-        return files
+        self.num_files = len(selected_files)
+        return selected_files
 
-    def collect_ply_data(self, ply_verts, target_array, array_postion):
+    def collect_ply_data(self, ply_verts, target_array, sample_number):
+        index = 0
         for vert in range(self._num_vertices):
             v = ply_verts[vert]
             (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
-            target_array[array_postion+vert] = [x, y, z]
+            target_array[index][sample_number] = x
+            target_array[index + 1][sample_number] = y
+            target_array[index + 2][sample_number] = z
+            index += 3
+    
+    def create_blendshapes(self, vertices, n_shapes):
+        # Subtract the mean
+        vertices = vertices - np.mean(vertices, axis=1, keepdims=True)
+        v = vertices.T @ vertices
 
-ply_files = "/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01"
+        # Eigen Analysis and sorting 
+        eigvals, eigvecs = la.eig(v)
[…]

### 2019-06-11 src/mesh/blendshape_interpolate.py
scripts to assist creating blendshapes

@@ -1,0 +1,38 @@
+import numpy as np
+import os 
+from plyfile import PlyData
+from blendshapes import Blendshapes
+from sklearn.preprocessing import normalize
+
+dir_path = os.path.dirname(os.path.realpath(__file__))
+ply_file = os.path.join(dir_path, '/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01/sentence01.000001.ply')
+shapes_file = os.path.join(dir_path, 'blendshapes_sentence01.txt')
+
+bs = Blendshapes(ply_files=ply_file)
+
+verts = np.empty((3 * bs._num_vertices, bs.num_files))
+
+with open(bs.ply_files, 'rb') as f:
+    plydata = PlyData.read(f)
+    bs.collect_ply_data(plydata['vertex'], verts)
+    faces = plydata['face'].data
+
+shapes = np.loadtxt(shapes_file, delimiter=',')
+shapes = normalize(shapes, axis=1)
+
+#princial_0 = verts + np.reshape(0.05 * shapes[:,0], verts.shape)
+#
+#bs.export_mesh(princial_0, faces, 'princ3')
+
+save_path = os.path.join(dir_path, 'blendshapes/shape00')
+
+shape = np.reshape(shapes[:,0], verts.shape)
+step = 0.001
+file_num = 0
+for i in range(-50, 51, 1):
+    filename = 'face' + '%05d' % file_num
+    path = os.path.join(save_path, filename)
+    weight = step * i
+    blend = verts + (weight * shape)
+    bs.export_mesh(blend, faces, path)
+    file_num += 1

### 2019-06-11 src/mesh/blendshapes.py

@@ -1,0 +1,114 @@
+from plyfile import PlyData, PlyElement
+import scipy.linalg as la
+import os
+import numpy as np
+from random import shuffle
+
+class Blendshapes():
+    def __init__(self, ply_files=None):
+        self.ply_files = ply_files
+        self._num_vertices = 5023
+        self.num_files = 1
+
+    def list_of_plys(self, keep=1):
+        """
+        Returns a list of ply files in given directory
+        If keep is less than 1, that proportion of files are kept.
+        """
+        selected_files = []
+        for root, _dirs, files in os.walk(self.ply_files):
+            for file in files:
+                if file.endswith(".ply"):
+                    selected_files.append(os.path.join(root, file))
+        
+        self.num_files = int(round(keep * len(selected_files)))
+        shuffle(selected_files)
+        del selected_files[self.num_files:]
+        
+        return selected_files
+
+    def collect_ply_data(self, ply_verts, target_array, sample_number=0):
+        index = 0
+        for vert in range(self._num_vertices):
+            v = ply_verts[vert]
+            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
+            target_array[index][sample_number] = x
+            target_array[index + 1][sample_number] = y
+            target_array[index + 2][sample_number] = z
+            index += 3
+    
+    def collect_mesh_faces(self, ply_file):
+        with open(ply_file, 'rb') as f:
+            plydata = PlyData.read(f)
+            faces = plydata['face'].data
+        return faces
+    
+    def create_blendshapes(self, vertices, n_shapes):
+        # Subtract the mean
+        vertices = vertices - np.mean(vertices, axis=1, keepdims=True)
+        v = vertices.T @ vertices
+
+        # Eigen Analysis and sorting 
+        eigvals, eigvecs = la.eig(v)
+        idx = eigvals.argsort()[::-1]   
+        eigvals = eigvals[idx]
+        eigvecs = eigvecs[:,idx]
+
+        eigvals_diag = np.diag(np.diag(eigvecs))
+
+        transform = vertices @ eigvecs @ la.inv(eigvals_diag)
+        blendshapes = np.delete(transform, np.s_[n_shapes:], axis=1)
+
+        return blendshapes
+    
+    def export_mesh(self, verts, faces, filename=None, text=False):
+        if filename == None:
[…]

### 2019-06-11 src/mesh/hello_mesh.py

@@ -1,43 +1,0 @@
-from plyfile import PlyData
-import os
-import numpy as np
-import time
-
-filename = "head.ply"
-dir_path = os.path.dirname(os.path.realpath(__file__))
-file_path = os.path.join(dir_path, filename)
-ply_files = "/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01"
-
-def ply_verts_to_array(ply_verts):
-    length = ply_verts.count
-    verts = np.empty((length, 3))
-
-    for vert in range(length):
-        v = ply_verts[vert]
-        (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
-        verts[vert] = [x, y, z]
-
-    return verts
-
-def list_of_plys(filepath):
-    """
-    Returns a list of ply files in given directory
-    """
-    files = []
-    for file in os.listdir(filepath):
-        if file.endswith("1.ply"):
-            files.append(file)
-    return files
-
-verts = []
-
-files = list_of_plys(ply_files)
-for file in files:
-    filepath = os.path.join(ply_files, file)
-
-    with open(filepath, 'rb') as f:
-        plydata = PlyData.read(f)
-        face_verts = ply_verts_to_array(plydata['vertex'])
-    verts.append(face_verts)
-
-print(len(verts))

### 2019-06-11 src/mesh/pca_mesh.py

@@ -1,72 +1,0 @@
-from plyfile import PlyData
-import scipy.linalg as la
-import os
-import numpy as np
-import time
-
-class Blendshapes():
-    def __init__(self, ply_files):
-        self.ply_files = ply_files
-        self._num_vertices = 5023
-        self.num_files = 0
-
-    def list_of_plys(self):
-        """
-        Returns a list of ply files in given directory
-        """
-        selected_files = []
-        for root, _dirs, files in os.walk(self.ply_files):
-            for file in files:
-                if file.endswith("1.ply"):
-                    selected_files.append(os.path.join(root, file))
-        
-        self.num_files = len(selected_files)
-        return selected_files
-
-    def collect_ply_data(self, ply_verts, target_array, sample_number):
-        index = 0
-        for vert in range(self._num_vertices):
-            v = ply_verts[vert]
-            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
-            target_array[index][sample_number] = x
-            target_array[index + 1][sample_number] = y
-            target_array[index + 2][sample_number] = z
-            index += 3
-    
-    def create_blendshapes(self, vertices, n_shapes):
-        # Subtract the mean
-        vertices = vertices - np.mean(vertices, axis=1, keepdims=True)
-        v = vertices.T @ vertices
-
-        # Eigen Analysis and sorting 
-        eigvals, eigvecs = la.eig(v)
-        idx = eigvals.argsort()[::-1]   
-        eigvals = eigvals[idx]
-        eigvecs = eigvecs[:,idx]
-
-        eigvals_diag = np.diag(np.diag(eigvecs))
-
-        transform = vertices @ eigvecs @ la.inv(eigvals_diag)
-        blendshapes = np.delete(transform, np.s_[n_shapes:], axis=1)
-
-        return blendshapes
-
-
-ply_files = "/home/peter/Documents/Uni/Project/datasets/registereddata/"#FaceTalk_170725_00137_TA"#/sentence01"
-
-bs = Blendshapes(ply_files)
-
-files = bs.list_of_plys()
-
-verts = np.zeros((3 * bs._num_vertices, bs.num_files))
-sample_number = 0
-for file in files:
-    filepath = os.path.join(ply_files, file)
-
-    with open(filepath, 'rb') as f:
-        plydata = PlyData.read(f)
-        bs.collect_ply_data(plydata['vertex'], verts, sample_number)
[…]

### 2019-06-11 src/mesh/blendshapes.py
reformatted a lot of blendshape/mesh code

@@ -1,114 +1,0 @@
-from plyfile import PlyData, PlyElement
-import scipy.linalg as la
-import os
-import numpy as np
-from random import shuffle
-
-class Blendshapes():
-    def __init__(self, ply_files=None):
-        self.ply_files = ply_files
-        self._num_vertices = 5023
-        self.num_files = 1
-
-    def list_of_plys(self, keep=1):
-        """
-        Returns a list of ply files in given directory
-        If keep is less than 1, that proportion of files are kept.
-        """
-        selected_files = []
-        for root, _dirs, files in os.walk(self.ply_files):
-            for file in files:
-                if file.endswith(".ply"):
-                    selected_files.append(os.path.join(root, file))
-        
-        self.num_files = int(round(keep * len(selected_files)))
-        shuffle(selected_files)
-        del selected_files[self.num_files:]
-        
-        return selected_files
-
-    def collect_ply_data(self, ply_verts, target_array, sample_number=0):
-        index = 0
-        for vert in range(self._num_vertices):
-            v = ply_verts[vert]
-            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
-            target_array[index][sample_number] = x
-            target_array[index + 1][sample_number] = y
-            target_array[index + 2][sample_number] = z
-            index += 3
-    
-    def collect_mesh_faces(self, ply_file):
-        with open(ply_file, 'rb') as f:
-            plydata = PlyData.read(f)
-            faces = plydata['face'].data
-        return faces
-    
-    def create_blendshapes(self, vertices, n_shapes):
-        # Subtract the mean
-        vertices = vertices - np.mean(vertices, axis=1, keepdims=True)
-        v = vertices.T @ vertices
-
-        # Eigen Analysis and sorting 
-        eigvals, eigvecs = la.eig(v)
-        idx = eigvals.argsort()[::-1]   
-        eigvals = eigvals[idx]
-        eigvecs = eigvecs[:,idx]
-
-        eigvals_diag = np.diag(np.diag(eigvecs))
-
-        transform = vertices @ eigvecs @ la.inv(eigvals_diag)
-        blendshapes = np.delete(transform, np.s_[n_shapes:], axis=1)
-
-        return blendshapes
-    
-    def export_mesh(self, verts, faces, filename=None, text=False):
-        if filename == None:
[…]

### 2019-06-11 src/mesh/mesh.py

@@ -1,0 +1,107 @@
+from plyfile import PlyData, PlyElement
+import scipy.linalg as la
+import os
+import numpy as np
+from random import shuffle
+
+class Mesh():
+    def __init__(self, dir_plys):
+        self.dir_plys = dir_plys
+
+        self.ply_files = []
+
+        for root, _dirs, files in os.walk(self.dir_plys):
+            for file in files:
+                if file.endswith(".ply"):
+                    self.ply_files.append(os.path.join(root, file))
+        
+        self.ply_files = sorted(self.ply_files)
+
+        with open(self.ply_files.pop(0), 'rb') as f:
+            plydata = PlyData.read(f)
+            self.vertex_count = plydata['vertex'].count
+            self.first_frame_vertex = np.empty((3*self.vertex_count, 1))
+            self._get_vertex_postions(plydata, self.first_frame_vertex)
+            self.first_frame_faces = plydata['face'].data
+
+        self.num_files = len(self.ply_files)
+        self.mesh_vertices = np.empty((3*self.vertex_count, self.num_files))
+
+    def _get_vertex_postions(self, plydata, target_array, file_number=0):
+        index = 0
+        for vert in range(self.vertex_count):
+            v = plydata['vertex'][vert]
+            (x, y, z) = (v[t] for t in ('x', 'y', 'z'))
+            target_array[index][file_number] = x
+            target_array[index+1][file_number] = y
+            target_array[index+2][file_number] = z
+            index += 3
+
+    def collect_vertex_postions(self, target_array):
+        """
+        Extracts vertex postions for all ply files
+        """
+        print("Number of files being processed: {}".format(self.num_files))
+
+        for file_number, file in enumerate(self.ply_files):
+            with open(file, 'rb') as f:
+                plydata = PlyData.read(f)
+                self._get_vertex_postions(plydata, target_array, file_number)
+                if (file_number % 10) == 0:
+                    print("Processing file number: {}".format(file_number))
+    
+    def export_mesh(self, mesh_vertices, filename=None, text=False):
+        """
+        Given numpy array of xyz postions for each vertex, export mesh back
+        to a ply file with the vertex connections from original imported files.
+        """
+        if filename == None:
+            filename = 'my_plyfile'
+
+        vertices = np.empty((self.vertex_count,), 
[…]

### 2019-06-18 Report/pca.tex
progress on procrustes write up

--- 2019-06-06 Report/Progress_Report/Background.tex
@@
 
-% some macros
-\input{../notation}
+\begin{document}
 
-\usepackage[nottoc,notlot,notlof]{tocbibind}
-\usepackage[T1]{fontenc}
-\usepackage{hyperref}
-\usepackage{indentfirst}
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{geometry}
- \geometry
- {
-     a4paper,
-     left=25mm,
-     right=25mm,
-     top=30mm,
-     bottom=30mm,
- }
+\section{Blendshapes and PCA}
 
-\title{MSc Background Report}
-\author{Peter Robertson}
-\date{}
-
-\begin{document}
-\maketitle
-
-% page numbering etc.
-\pagenumbering{roman}
-\setcounter{page}{0}
-\clearpage{\pagestyle{empty}\cleardoublepage}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%--- table of contents
-\tableofcontents 
-\clearpage{\pagestyle{empty}\cleardoublepage}
-
-\pagenumbering{arabic}
-\setcounter{page}{1}
-
-\section{Introduction}
-
-\textbf{( I want to rewrite this )}
-
-Current methods of visual speech prediction, often referred to as lip reading, focuses on training machine learning models on 2D temporal data of participants speaking, labelled with the accompanying spoken text \cite{Chung2016, Assael2016, Chung2017, Shillingford2018}.
-However there has been little progress in lip reading from 3D temporal models which contain depth information of the subject.
-A substantial barrier to this problem is the lack of 3D datasets of subjects speaking when compared to the relative abundance of video datasets.
-To first attempt this problem, new datasets which capture 3D temporal models of subjects heads speaking must be established by the community.
-
-\section{Lip Reading}
-This section shall discuss current lip reading models built with deep learning techniques and the availability of both 2D and 3D temporal datasets which lip reading models can currently be trained on.
-In recent years the problem of Lip Reading has seen huge advances due to the availability of new datasets and the use of Deep Learning models.
-Due to the large availability of video data, datasets such as GRID \cite{Cooke2006}, LRW \cite{Chung2016}, LRS \cite{Chung2017} and LSVSR \cite{Shillingford2018} have been constructed by compiling the appropriate data.
-
-Current lip reading models currently make use of 2D temporal data such as videos.
-This is convenient as video data is already widely available and relatively straightforward to obtain and process.
-Unlike 2D temporal data, 3D temporal data (3D video), requires multiple cameras to record simultaneously, which requires synchronisation, adding to the complexity of the system.
-The data must then be processed to produce the final product, whether this is in the form of video with depth information or more complex 3D scans \cite{Li2017}.
-
-\subsection{Datasets}
-
-\subsubsection{Controlled Conditions Datasets}
-The GRID dataset was released in 2006 \cite{Cooke2006} and contains 34,000 samples from 34 speakers, each with 1000 sentences.
-The corpus uses sentences with a fixed grammar: 
-<command:4>, <colour:4>, <preposition:4>, <letter:25>, <number:10>, <adverb:4>, with a total vocabulary of 51 words.
-The primary limiting factor of the GRID dataset is that all data was captured directly for the use of the dataset.
-
-\subsubsection{In The Wild Datasets}
-To build larger datasets, subsequent datasets are commonly built with data "in the wild", meaning that the videos have not been captured with the intention of being used for this dataset.
-Variations in lighting, angles, speakers and a wide vocabulary are common, this does however make the datasets more challenging to learn from.
-LRW and LRS are both comprised of content from BBC broadcasts \cite{Chung2016, Chung2017} allowing for a far larger corpus size of over 1000 words for LRW and over 6000 words for LRS.
-Both of datasets are captured and processed with the same pipeline summarised as follows. 
-Firstly, as the subtitles are not aligned to the video on the broadcasts, optical chapter recognition is used to obtain the text being spoken in the video clips, the audio and text are then aligned per frame using the HTK toolkit \cite{Woodland1995}.
-A HOG-based detection algorithm \cite{King2009} is used for face detection for cropping the frames to the subjects head followed by facial landmarks for mouth localisation and speaker identification.
-
-LSVSR is a dataset published by DeepMind and Google which makes use of the huge amount of videos on YouTube \cite{Shillingford2018}, resulting in a total length of 3886 hours of training data.
-The dataset is far larger than LRS, but aligns phonemes to frames, as opposed to words or characters.
-The pre-processing steps are similar as in LRW and LRS; alignment is performed with the algorithm laid out in previous work by DeepMind \cite{Liao2013}, faces are tracked to ensure the speaker is visible in frame.
-
-\subsection{3D Datasets} \label{3D Datasets}
-To the best of the author's knowledge, there are currently two datasets with 3D temporal data which are appropriate for training lip reading models.
-The first of which is LRW-3D \cite{Tzirakis2019} which has been captured from four subjects, two native English speakers and two non-native to increase variability in the dataset.
-The subjects have been captured speaking the corpus used in the LRW dataset \cite{Chung2016}, a vocabulary of 500 words.
-The resulting dataset comprises of 660 seconds of 3D meshes and audio per subject.
-The dataset is not comprised of full sentences but would be appropriate for word-level lip reading.
-
-The second dataset is the VOCASET \cite{Cudeiro2019}, captured from 6 male and 6 female subjects.
-Each subject was recorded speaking 40 sequences, each ranging from 3 to 5 seconds, resulting in a total time of 30 minutes.
-The recorded 3D meshes are registered to the FLAME model \cite{Li2017}, a statistical 3D facial mesh with around 5000 vertices.
-Unlike the LRW-3D dataset, the sequences are grammatically correct sentences, chosen to maximise phonetic diversity.
-This makes the VOCASET appropriate for creating a model for sentence-level lip reading.
-
-It should be noted that neither of these datasets were captured for the purpose of lip reading, but for synthesising realistic statistical facial models driven from an audio input, and thus the transcriptions and frames are not aligned.
-This alignment would have to be performed to the datasets before using the data to train lip reading models.
-In order to achieve this automatic speech recognition (ASR) systems such as DeepSpeech \cite{Hannun2014} could be used to solve this problem.
-
-\subsection{Lip Reading Models}
-Chung et al \cite{Chung2016} used a convolutional model based on the VGG-M architecture which produced character level distributions with the LRW dataset, these distributions are then processed by a language model for text prediction.
-
-The LipNet model \cite{Assael2016} was produced to be able to predict sentence-level lip reading of varied length, while previous work by \cite{Chung2016} predicted on a word level.
-The model used spatiotemporal convolutions to process multiple frames of video at once followed by a recurrent layer using Gated Recurrent Units (GRU) \cite{Cho2014}.
-The model used the GRID dataset \cite{Cooke2006} in which all subjects are recorded under consistent conditions of good lighting and fixed angles.
-LipNet achieved the state of the art performance on the GRID dataset by using an model with increased complexity by incorporating 3D convolution and recurrent units.
-
-As the LRW dataset could not be used to train a model such as LipNet for sentence-level lip reading, but the GRID dataset had limitations in the number of subjects and vocabulary, Chung et al created the LRS dataset \cite{Chung2017}.
-The model presented was trained on both audio and video and made capable of taking either or both as the model inputs.
-To prevent the model from being dependent on a single input source, the inputs are systematically distorted or removed.
-The video input is passed through convolutional layers, followed by LSTM layers, while the audio is converted to MFCC, then input to LSTM layers.
-The two are combined with then attention mechanism and further LSTM layers and an output fully connected layer with softmax activation for character distributions.
-The model also makes use of curriculum learning by initially training the model on short sequences of single words.
-The length of training sequences are increased throughout training.
-It is stated by \cite{Chung2017} that this accelerates training and reduces overfitting.
-
-In 2018 DeepMind published their V2P model \cite{Shillingford2018} along with the LSVSR dataset which is larger than all previous datasets, containing 3886 hours of training data.
-The model follows a similar architecture to LipNet \cite{Assael2016} but with an increased number of convolutional layers and LSTM layers rather than GRUs.
-It should be commented that due to the size of the model and dataset dictated the use of 64 GPUs for training to allow a batch size of 128.
-Unlike previous models, the V2P model predicts phonemes as opposed to characters, these phonemes are then processed by a language model for word prediction as with previous models.
-
-To the best of the author's knowledge, currently there do not exist any papers which have explored the use of 3D temporal datasets for the use with lip reading models. 
-This is likely due to the shortage of 3D temporal data on which such models could be trained on.
-
-\section{Data Generation}
-In order to construct a deep learning lip reading model capable of being trained on 3D temporal data, appropriate datasets must be established.
-Current datasets have been captured directly \cite{Tzirakis2019, Cudeiro2019} with the use of multi-camera capture rigs under controlled situations.
-The total duration of both of these datasets is very short in comparison to the video datasets such as LRW, LRS and LSVSR, which is a limiting factor as to the models which could be trained using them.
-As the models also use different mesh models to represent the data that has been captured this also prevents the two datasets being joined directly.
-Unlike video data, there currently lacks a large body of 3D video data which is publicly available, limiting the construction of 3D datasets to directly capturing more 3D scans with multi-camera capture rigs, similar to those used in \cite{Tzirakis2019, Cudeiro2019} and generating synthetic training data.
-
-\subsection{Audio Driven Data Generation}
-
-\subsubsection{Imperial}
-\cite{Tzirakis2019}
-
-\subsubsection{VOCA}
-The VOCA model \cite{Cudeiro2019} synthesises video sequences of 3D models speaking given an audio input.
-The VOCASET dataset discussed in section \ref{3D Datasets}, was captured with the intention of training this model to be independent of the speaker, hence a large range of speakers are used within the dataset.
-The model is comprised of three sections: audio feature extraction, a feature encoder and a decoder to drive a template facial mesh from the FLAME model \cite{Li2017}.
-The audio feature extraction makes use of the pre-trained Mozilla implementation of the DeepSpeech model, based on the paper by Hannun et al. \cite{Hannun2014}.
-The DeepSpeech model takes audio as an input and returns the unnormalised log-probabilities for an alphabet of the 26 standard characters, a space, apostrophe and blank character for time slices in the audio input.
-The encoder is a convolutional network which is conditioned on the speakers identity, such that the latent space of speaker styles can later be explored on new audio inputs.
-Finally, the decoder is made up of a fully connected layer with a linear activation function is used to output the displacements of the 5023 vertices in the template face.
-
-\subsubsection{Nvidia}
-\cite{Karras2017a}
-
-\subsection{Adversarial Trained Generative Methods}
-A recent development in generating synthetic data samples is the use of generative adversarial networks \cite{Goodfellow2014}.
-The concept behind generative adversarial networks (GANs) is two have two machine learning models; a generator and a discriminator.
-The task of the generator is to produce samples from an unknown high order probability distribution which correctly resemble samples from the distribution defined by training data.
-The generator achieves this by transforming a random sample from a known probability distribution, such as a Gaussian distribution as used in \cite{Goodfellow2014}, to a sample from the unknown distribution.
-This is achieved by finding the function which maps between the two distributions.
-The discriminator however, attempts to correctly learn to discriminate between the real and the fake generated samples.
-
-\begin{equation} \label{eq:gans_loss}
-    \min_{G} \max_{D} V(G, D) = \E{\bm{x} \sim p_{data}(\bm{x})} [\log D(\bm{x})]
-                              + \E{\bm{z} \sim p_{z}(\bm{z})} [\log (1 - D(G(\bm{z})))]
-\end{equation}
-
-\noindent The original loss function (\ref{eq:gans_loss}) proposed by Ian Goodfellow forms a min-max game, where the loss of the generator is attempting to be minimised by having the discriminator label all the generated samples as real.
-While the loss of the discriminator is maximised by correctly classifying real and fake samples.
-The two networks are trained in an alternating fashion until the discriminator achieves an accuracy of 50\%, effectively making binary guesses between real and generated samples.
-
-GANs however, are difficult to train for two main reasons.
-Firstly, the equation (\ref{eq:gans_loss}) is challenging as it provides small gradients while generated samples are poor as discussed in \cite{Goodfellow2014}, making training difficult.
-Progress in developing new loss functions is discussed in section \ref{Stability_to_GANs}
-Secondly, early networks must also be balanced with a similar model capacity to prevent one from getting too much better than the other, preventing the other from improving.
-Various architectural changes have improved this issue \cite{Radford2016, Zhang2018}, although it seems to be closely tied to the loss function being used \cite{Gulrajani2017}.
-
-\subsubsection{Stability Improvements of GANs Networks} \label{Stability_to_GANs}
-There have been a large number of papers presenting new techniques with differing levels of success and training stability, a small handful of key papers which provided large advances in the generative adversarial training model shall be discussed here.
-A primary research focus around GANs has been in finding new loss functions on which to train the model to improve stability and performance.
-The original loss function proposed in the original paper \cite{Goodfellow2014} identifies an issue with equation (\ref{eq:gans_loss}) in that the gradient back propagated to the generator when the generated samples are poor, is very low as shown in figure \ref{fig:Goodfellow_plot}.
-
-\begin{figure}[h]
-    \centering
-        \includegraphics[width=0.8\textwidth]{../figures/goodfellow_gen_losses.png}
-    \caption{Goodfellow's Generator Loss Plots \cite{Goodfellow2014}}\label{fig:Goodfellow_plot}
-\end{figure}
-
-\noindent This in turn makes it very challenging to improve the performance of the generator.
-The suggested improvement made in \cite{Goodfellow2014} is rather than to maximise the the number of generated samples which the discriminator incorrectly classifies, but to minimise the number of generator samples the discriminator correctly classifies, described in equation (\ref{eq:gans_loss2}).
-Figure \ref{fig:Goodfellow_plot} shows how this results in the gradient propagated back to the generator is far larger when the generated samples are poor.
-
-\begin{equation} \label{eq:gans_loss2}
-    \min_{G} \max_{D} V(G, D) = \E{\bm{x} \sim p_{data}(\bm{x})} [\log D(\bm{x})]
-                              - \E{\bm{z} \sim p_{z}(\bm{z})} [\log (D(G(\bm{z})))]
-\end{equation}
-
-Further stability improvements were proposed in \cite{Radford2016} which allowed deep convolutional generative adversarial networks (DCGANs) to be successfully trained for the first time.
-Radford et al. proposed three main contributions.
-Replace deterministic pooling layers with strided convolutions in both the generator and discriminator networks, allowing the networks to learn their own spatial upsampling and downsampling.
-Remove all fully connected layers used on top of convolutional layers, resulting in a fully convolutional model.
-Apply batch normalisation \cite{Ioffe2015} before the input of each layer.
-This normalises the input to each unit to zero mean and unit variance.
-This assists with training problems due to poor weight initialisation and allows gradients to flow through deeper networks more easily.
-Radford et al. state this to be a critical improvement to allow generator networks to begin learning by preventing all samples from collapsing to a single point.
-
-In an attempt to stabilize the training of GANs models, the Wasserstein or 'Earth Mover Distance' loss function was proposed by Arjovsky et al.
-\cite{Arjovsky2017} shown in equation (\ref{eq:wgan}) where $\mathcal{D}$ is a set of 1-Lipschtiz functions.
-
-\begin{equation} \label{eq:wgan}
-    \min_{G} \max_{D \in \mathcal{D}} W(\Pd{r}, \Pd{g}) =
-            \E{\bm{x} \sim \Pd{r}} [D(\bm{x})]
-            -\E{\bm{\hat{x}} \sim \Pd{g}} [D(\hat{\bm{x}})]
-\end{equation}
-
-\noindent The WGAN model uses a critic as opposed to a discriminator, this is due to the fact that the discriminator is no longer a binary classifier, but being used to critique the real and generated samples.
-As the Wasserstein function is continuous and differentiable, the critic can be trained until optimality, and \cite{Arjovsky2017} argues that it should be.
-As the critic is trained to optimality it does not saturate, but converges to a linear function.
-This provides the generator with a gradient which is more reliable, resulting in the generator learning consistently.
-The Wasserstein loss function has been shown to greatly improve training stability by providing consistent gradients throughout training and no longer requiring that the two networks have a balanced model capacity.
-
-In order to use the Wasserstein distance as the loss for WGAN, the Lipschtiz condition must be enforced.
-In the WGAN the model weights are clipped to enforce this constraint \cite{Arjovsky2017}, however this is stated to be a non-ideal method of achieving this constraint and calls for future work to be investigate more effective methods.
-The use of gradient penalty is proposed in \cite{Gulrajani2017} in order to satisfy this condition more elegantly as described in equation (\ref{eq:wgan_gp}).
-
-\begin{equation} \label{eq:wgan_gp}
-    \min_{G} \max_{D \in \mathcal{D}} W(\Pd{r}, \Pd{g}) 
-        = \E{\bm{x} \sim \Pd{r}} [D(\bm{x})]
-        - \E{\bm{\tilde{x}} \sim \Pd{g}} [D(\tilde{{\bm{x}}})]
-        + \lambda \E{\bm{\hat{x}} \sim \Pd{\bm{\hat{x}}}} 
-            [(\| \nabla_{\bm{\hat{x}}} D(\bm{\hat{x}}) \| - 1)^2]
-\end{equation}
-
-\noindent Using the Wasserstein loss function with gradient penalty enforces the Lipschtiz condition and allows highly complex architectures to be trained successfully, including those with residual units \cite{Gulrajani2017}.
-
-\subsubsection{Architectural Developments}
-Mirza et al. showed that GANs could be conditioned on additional inputs to the generator network as the noise component \cite{Mirza2014}.
-This has allowed facilitated other uses of GANs such as image to image translation for style transfer \cite{Zhu2017}.
-Vougioukas et al.
-used a temporal model to generate a sequence of video frames given an image of a subject and an audio sequence of spoken text to synthesise the subject speaking \cite{Vougioukas2018}.
-The model uses two discriminators, one to determine if individual frames are realistic images of the subject's face and a second which evaluates the sequence of video frames to determine if it is realistic.
-The model uses temporal components to examine if the frames of video are consistent in time, preventing sudden jumps in facial position.
-
-Other novel architectures include the progressively growing GAN model \cite{Karras2017b} which is able to generate highly realistic images of faces to a high resolution.
-This is achieved by initially training a shallow model to produce 4x4 pixel images before increasing the depth of the network and training further at a higher resolution.
-By forcing the model to firstly produce and examine low resolution images the model firstly has to be able to synthesise simple low level features effectively, such as facial shape which are common to all samples.
-Once the model can produce these features a higher resolution is used, allowing it to learn more complex features, such as hair and eyes.
-
-The attention mechanism \cite{Vaswani2017} has been shown to be usefully when applied to image data \cite{Xu2015} in order to capture relationships between spatially distant points.
-As such points are further apart in the image, previously deeper convolutional models were required to allow a large enough receptive field to capture information on these two points.
-Zhang et al.
-point out that this is a common issue with DCGANs \cite{Radford2016}, where generated samples fail to produce structural patterns, while they do exceedingly well at local textural patterns.
-This is often seen in generated images of animals with realistic fur, but oddly shaped or an incorrect number of limbs.
-The attention mechanism is applied to GANs \cite{Zhang2018} to allow for long range dependencies in the image to be modelled by convolutional models more effectively.
-
-\section{3D Modelling}
-\begin{itemize}
-    \item blendshapes
-    \item Model Correspondence
-\end{itemize}
-
-\section{Recurrent Models}
-\begin{itemize}
-    \item Recurrent networks
-    \item LSTMs
-    \item Attention Mechanism
-\end{itemize}
-
-\section{Speech Recognition}
-Deep Speech \cite{Karras2017a}
-
-\section{Current Work and Future Work}
-Data collection
-
-\bibliographystyle{unsrt}
-\bibliography{../ref}
 \end{document}

### 2019-06-18 Report/procrustes.tex
progress on procrustes write up

\documentclass[12pt]{article}
% some macros
\input{notation}
\input{includes}

\begin{document}

\section{Procrustes for mesh alignment}
In order to create blendshapes from 3D head scans, movement within scans should be limited to facial movement with as little head movement as possible.
If head movement remains within the scans, then this will be reflected in the principal axis produced by PCA.
As head motion is independent from speech, this is undesirable.
Initial data capture can aim to minimise subject head movement, however this cannot be completely eliminated.
To remove head motion from the dataset, the scans can be brought into alignment relative to landmark positions on the face which do not move during speech.
Such landmarks include the nose, corners of the eyes and sides of the head, as these points which will be stationary during speech.
By aligning the meshes based on the variation in these points, the head position of the subjects can be constrained while the mouths are not.
To achieve this, Procrustes Analysis can be used.

\subsection{Procrustes Analysis}
Procrustes Analysis is a statistical shape analysis method used to analyse the differences between two objects \cite{Krzanowski2000}
Given matrices $\mat{X} \in \mathbb{R}^{(n\times f)}$ and $\mat{Y} \in \mathbb{R}^{(n\times f)}$, each representing the coordinates of $n$ points in an $f$-dimensional feature space.
A common statistical difference metric is the sum of square differences \ref{eq:ssd}.
\begin{equation}
    \label{eq:ssd}
    D = \sum_{i=1}^{n} \sum_{j=1}^{f} (x_{ij} - y_{ij})^2
\end{equation}

\noindent However, two objects which are mathematically similar can still have a significant difference when they are scaled differently and are in different positions, orientations in space.
In order to compare two objects regardless of these factors, the differences in orientation, scale and spacial position of the objects must be minimised by bringing the objects into optimal alignment.

\subsubsection{Translational Alignment}
Translational components can be eliminated by having the centroids of the two objects lie at the same position in space.
This can be easily achieved by subtracting the mean of each objects points from itself such that the centroid now lies at the origin.

Let $\bar{x}_j = \frac{1}{n} \sum_{i=1}^{n} x_{ij}$ and $\bar{y}_j = \frac{1}{n} \sum_{i=1}^{n} y_{ij}$ where $(j = 1, \dots, f)$, such that the centroids of the two objects are given by $C_X = (\bar{x}_1, \bar{x}_2, \dots, \bar{x}_f)$ and $C_Y = (\bar{y}_1, \bar{y}_2, \dots, \bar{y}_f)$ respectively.
By subtracting the mean of $\mat{X}$ and $\mat{Y}$ from themselves, the centroids $C_X$ and $C_Y$ will be in alignment at the origin, minimising the sum of square differences due to translational components.

\subsubsection{Scaling Alignment}
Similarly, scaling components can be removed by rescaling the objects by the root mean square distance (\ref{eq:rms}), such that the root mean square distance will be of unit distance for both.

\begin{equation}
    \label{eq:rms}
    D = \sqrt{\frac{1}{n} \sum_{i=1}^{n} \sum_{j=1}^{f} x_{ij}^2}
\end{equation}

\subsubsection{Rotational Alignment}
In order to align the objects in the same orientation, a rotational matrix which is able to map $\mat{X}$ to $\mat{Y}$ as closely as possible must be found.
This is described as the Orthogonal Procrustes problem and can be solved with singular value decomposition (SVD).
Given matrices $\mat{A}$ and $\mat{B}$, the orthogonal matrix $\mat{R}$ is the matrix which matches $\mat{A}$ to $\mat{B}$ as closely as possible, as in equation (\ref{eq:orth_proc_prob}).

\begin{equation}
    \label{eq:orth_proc_prob}
    \mat{R} =\operatorname*{arg}\operatorname*{min}_\mat{\Omega} 
        \| \mat{\Omega} \mat{A} - \mat{B} \|_F
        \quad
        \text{subject to}
        \quad
        \mat{\Omega}^\top \mat{\Omega} = \mat{I}
\end{equation}

\noindent
The product of matrices $\mat{A}$ and $\mat{B}$ can be decomposed (\ref{eq:svd_2}), and then the rotational matrix $\mat{R}$ can be expressed (\ref{eq:proc_rotation}), where $\| \cdot \|_F$ is the Fobius norm.
\begin{equation} \label{eq:svd_1}
    \mat{M} = \mat{B} \mat{A}^\top
\end{equation}
\begin{equation} \label{eq:svd_2}
    \mat{M} = \mat{U} \mat{\Sigma} \mat{V}^\top
\end{equation}
\begin{equation} \label{eq:proc_rotation}
    \mat{R} = \mat{U} \mat{V}^\top
\end{equation}

\subsection{A Simple Application of Procrustes Analysis}
This section follows the steps of object alignment using Procrustes Analysis as described above with a simple example.
A pair of configurations are given by the matrices below, visualised in figure \ref{fig:procrustes_ex_1}.

\begin{equation*}
    \mat{X} = 
    \begin{pmatrix} 
        1 & 1 \\
        1 & 2 \\
        3 & 2
    \end{pmatrix}
    \quad 
    \text{and} 
    \quad 
    \mat{Y}
    \begin{pmatrix} 
        -3 & -2 \\
        -3 & -4 \\
        -7 & -4
    \end{pmatrix}
\end{equation*}

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.475\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/procrustes_ex1}
        \caption[]
        {{\small Original coordinates}}    
        \label{fig:procrustes_ex_1}
    \end{subfigure}
    \begin{subfigure}[b]{0.475\textwidth}  
        \centering 
        \includegraphics[width=\textwidth]{figures/procrustes_ex2}
        \caption[]
        {{\small Translation}}    
        \label{fig:procrustes_ex_2}
    \end{subfigure}
    \begin{subfigure}[b]{0.475\textwidth}   
        \centering 
        \includegraphics[width=\textwidth]{figures/procrustes_ex3}
        \caption[]
        {{\small Scaling}}    
        \label{fig:procrustes_ex_3}
    \end{subfigure}
    \begin{subfigure}[b]{0.475\textwidth}   
        \centering 
        \includegraphics[width=\textwidth]{figures/procrustes_ex4}
        \caption[]
        {{\small Rotation}}    
        \label{fig:procrustes_ex_4}
    \end{subfigure}
    \caption[]
    {Application of Procrustes Analysis} 
    \label{fig:procrustes_analysis}
\end{figure*}

\noindent
Before applying alignment, the sum of square differences between $\mat{X}$ and $\mat{Y}$ is
\begin{align*}
    D& = {(1+3)^2 + (1+2)^2} + {(1+3)^2 + (2+4)^2} + {(3+7)^2 + (2+4)^2} \\
     & = 213
\end{align*}
while after the alignment, the difference is zero as these objects are both similar triangles.

\newpage
\bibliographystyle{unsrt}
\bibliography{ref}
\end{document}

### 2019-06-18 Scripts/procrustes_plot.py

@@ -1,0 +1,56 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.linalg as la
+
+x = np.array([[1,1],[1,2],[3,2]], dtype=np.float)
+y = -2*x - [1,0]
+print(x)
+print(y)
+c_x = np.mean(x, 0)
+c_y = np.mean(y, 0)
+plt.scatter(x[:,0], x[:,1], marker='o', label='X')
+plt.scatter(y[:,0], y[:,1], marker='x', label='Y')
+plt.scatter((c_x[0], c_y[0]),(c_x[1], c_y[1]), label='Centroids')
+plt.legend()
+plt.grid()
+plt.show()
+
+# Centre
+x_mean = np.mean(x, 0)
+y_mean = np.mean(y, 0)
+x -= x_mean
+y -= y_mean
+c_x = np.mean(x, 0)
+c_y = np.mean(y, 0)
+plt.scatter(x[:,0], x[:,1], marker='o', label='X')
+plt.scatter(y[:,0], y[:,1], marker='x', label='Y')
+plt.scatter((c_x[0], c_y[0]),(c_x[1], c_y[1]), label='Centroids')
+plt.legend()
+plt.grid()
+plt.show()
+
+# Scale
+norm_x = la.norm(x)
+norm_y = la.norm(y)
+x /= norm_x
+y /= norm_y
+c_x = np.mean(x, 0)
+c_y = np.mean(y, 0)
+plt.scatter(x[:,0], x[:,1], marker='o', label='X')
+plt.scatter(y[:,0], y[:,1], marker='x', label='Y')
+plt.scatter((c_x[0], c_y[0]),(c_x[1], c_y[1]), label='Centroids')
+plt.legend()
+plt.grid()
+plt.show()
+
+# Rotate
+rotation, _ = la.orthogonal_procrustes(x, y)
+y = (y @ rotation.T)
+c_x = np.mean(x, 0)
+c_y = np.mean(y, 0)
+plt.scatter(x[:,0], x[:,1], marker='o', label='X')
+plt.scatter(y[:,0], y[:,1], marker='x', label='Y')
+plt.scatter((c_x[0], c_y[0]),(c_x[1], c_y[1]), label='Centroids')
+plt.legend()
+plt.grid()
+plt.show()

### 2019-06-18 src/mesh/mesh.py

@@ -134,7 +134,7 @@ class Mesh():
         vertices = vertices[:,1::2] - vertices[:,::2]
         return vertices
     
-    def _get_procrustes_parameters(self, matrix1, matrix2):
+    def _apply_procrustes(self, matrix1, matrix2):
         # Translate centroids of matricies to origin
         mean1 = np.mean(matrix1, 0)
         mean2 = np.mean(matrix2, 0)
@@ -147,24 +147,9 @@ class Mesh():
         matrix1 /= norm1
         matrix2 /= norm2
 
-        rotation, scale = la.orthogonal_procrustes(matrix1, matrix2)
-
-        return mean1, mean2, norm1, norm2, rotation, scale
-    
-    def _apply_procrustes(self, matrix1, matrix2, params):
-        """
-        params = [mean_matrix1, 
-                  mean_matrix2, 
-                  norm_matrix1, 
-                  norm_matrix2,
-                  rotation_matrix,
-                  scaling]
-        """
-        matrix1 -= params[0]
-        matrix2 -= params[1]
-        matrix1 /= params[2]
-        matrix2 /= params[3]
-        matrix2 = (matrix2 @ params[4].T) * params[5]
+        # Rotate matrix 2
+        rotation, _ = la.orthogonal_procrustes(matrix1, matrix2)
+        matrix2 = (matrix2 @ rotation.T)
 
         return matrix1, matrix2
     
@@ -182,8 +167,7 @@ class Mesh():
             submat1[i] = matrix1[landmark_idx]
             submat2[i] = matrix2[landmark_idx]
         
-        params = self._get_procrustes_parameters(submat1, submat2)
-        matrix1, matrix2 = self._apply_procrustes(matrix1, matrix2, params)
+        matrix1, matrix2 = self._apply_procrustes(matrix1, matrix2)
 
         diff = np.sum(np.square(matrix1 - matrix2))
 

### 2019-06-19 Report/procrustes.tex
finished procrustes section. Would be nice to have an example of facial mesh alignment

@@
 
-\section{Procrustes for mesh alignment}
-In order to create blendshapes from 3D head scans, movement within scans should be limited to facial movement with as little head movement as possible.
-If head movement remains within the scans, then this will be reflected in the principal axis produced by PCA.
-As head motion is independent from speech, this is undesirable.
+\section{Facial Mesh Alignment}
+In order to create blendshapes from 3D facial scans, movement within scans should be limited to facial movement with as little head movement as possible.
+If head movement remains within the scans, then this will be reflected in the principal axis produced by Principle Components Analysis (PCA) and as head motion is independent from speech, this is undesirable.
 Initial data capture can aim to minimise subject head movement, however this cannot be completely eliminated.
 To remove head motion from the dataset, the scans can be brought into alignment relative to landmark positions on the face which do not move during speech.
-Such landmarks include the nose, corners of the eyes and sides of the head, as these points which will be stationary during speech.
+Such landmarks include the nose, corners of the eyes and cheek bones, as these points which will be stationary during speech.
 By aligning the meshes based on the variation in these points, the head position of the subjects can be constrained while the mouths are not.
@@
 
-\subsection{Procrustes Analysis}
-Procrustes Analysis is a statistical shape analysis method used to analyse the differences between two objects \cite{Krzanowski2000}
-Given matrices $\mat{X} \in \mathbb{R}^{(n\times f)}$ and $\mat{Y} \in \mathbb{R}^{(n\times f)}$, each representing the coordinates of $n$ points in an $f$-dimensional feature space.
-A common statistical difference metric is the sum of square differences \ref{eq:ssd}.
+\subsection{Procrustes Analysis} \label{sec:procrustes_analysis}
+Procrustes Analysis is a statistical shape analysis method used to analyse the differences between two objects \cite{Krzanowski2000}.
+Given matrices $\mat{X} \in \mathbb{R}^{(n\times f)}$ and $\mat{Y} \in \mathbb{R}^{(n\times f)}$, each representing the coordinates of $n$ points in an $f$-dimensional feature space, a common statistical difference metric is the sum of square differences \ref{eq:ssd}.
 \begin{equation}
@@
 
-\subsubsection{Translational Alignment}
+\subsubsection{Translational Alignment} \label{sec:trans_align}
 Translational components can be eliminated by having the centroids of the two objects lie at the same position in space.
@@
 
-Let $\bar{x}_j = \frac{1}{n} \sum_{i=1}^{n} x_{ij}$ and $\bar{y}_j = \frac{1}{n} \sum_{i=1}^{n} y_{ij}$ where $(j = 1, \dots, f)$, such that the centroids of the two objects are given by $C_X = (\bar{x}_1, \bar{x}_2, \dots, \bar{x}_f)$ and $C_Y = (\bar{y}_1, \bar{y}_2, \dots, \bar{y}_f)$ respectively.
+Let $\bar{x}_j = \frac{1}{n} \sum_{i=1}^{n} x_{ij}$ and $\bar{y}_j = \frac{1}{n} \sum_{i=1}^{n} y_{ij}$ where $(j = 1, \dots, f)$ represent the mean of each feature, such that the centroids of the two objects are given by $C_X = (\bar{x}_1, \bar{x}_2, \dots, \bar{x}_f)$ and $C_Y = (\bar{y}_1, \bar{y}_2, \dots, \bar{y}_f)$ respectively.
 By subtracting the mean of $\mat{X}$ and $\mat{Y}$ from themselves, the centroids $C_X$ and $C_Y$ will be in alignment at the origin, minimising the sum of square differences due to translational components.
 
-\subsubsection{Scaling Alignment}
+\subsubsection{Scaling Alignment} \label{sec:scale_align}
 Similarly, scaling components can be removed by rescaling the objects by the root mean square distance (\ref{eq:rms}), such that the root mean square distance will be of unit distance for both.
@@
 
-\subsubsection{Rotational Alignment}
+\subsubsection{Rotational Alignment} \label{sec:rot_align}
 In order to align the objects in the same orientation, a rotational matrix which is able to map $\mat{X}$ to $\mat{Y}$ as closely as possible must be found.
-This is described as the Orthogonal Procrustes problem and can be solved with singular value decomposition (SVD).
-Given matrices $\mat{A}$ and $\mat{B}$, the orthogonal matrix $\mat{R}$ is the matrix which matches $\mat{A}$ to $\mat{B}$ as closely as possible, as in equation (\ref{eq:orth_proc_prob}).
+This is described as the Orthogonal Procrustes problem and can be solved with Singular Value Decomposition (SVD).
+Given matrices $\mat{A}$ and $\mat{B}$, the orthogonal matrix $\mat{R}$ is the matrix which matches $\mat{A}$ to $\mat{B}$ as closely as possible, as described by equation (\ref{eq:orth_proc_prob}), where $\| \cdot \|_F$ is the Fobius norm.
 
@@
 \noindent
-The product of matrices $\mat{A}$ and $\mat{B}$ can be decomposed (\ref{eq:svd_2}), and then the rotational matrix $\mat{R}$ can be expressed (\ref{eq:proc_rotation}), where $\| \cdot \|_F$ is the Fobius norm.
+The product of matrices $\mat{A}$ and $\mat{B}$ can be decomposed (\ref{eq:svd_2}), and then the rotational matrix $\mat{R}$ can be expressed (\ref{eq:proc_rotation}).
 \begin{equation} \label{eq:svd_1}
@@
 
-\subsection{A Simple Application of Procrustes Analysis}
+\subsubsection{A Simple Application of Procrustes Analysis}
 This section follows the steps of object alignment using Procrustes Analysis as described above with a simple example.
@@
 \end{align*}
-while after the alignment, the difference is zero as these objects are both similar triangles.
+As described in section (\ref{sec:trans_align}), the initial objects are translated so that their centroids lie at the origin (Figure \ref{fig:procrustes_ex_2}).
+This is then followed by scaling alignment and finally rotational alignment (Figures \ref{fig:procrustes_ex_3}, \ref{fig:procrustes_ex_4}).
+After alignment, the difference is zero as these objects are both similar triangles.
+
+\subsection{Procrustes Analysis for Facial Mesh Alignment}
+The process of Procrustes Analysis described above deals with objects in which all coordinate points are used to find the optimum alignment, however this is inappropriate for facial mesh alignment as not only would the position of the head be aligned to, but as would the motion from speech, of which the aim is to maintain variation.
+However, a subset of points of a pair of objects can be used to align the entire object.
+By selecting points which do not contain any movement due to speech or facial expression, variation in these points can be assumed to be only from head motion.
+Such positions include the nose, corners of the eyes and cheekbones.
+If the facial scans capture 360 degrees, then points on the back of the subjects head can also be used.
+From this subset of points, the translation, scaling and rotational matrix can be found which aligns these points using the steps described in section \ref{sec:procrustes_analysis} but then applied to the entire object.
 

### 2019-06-20 Report/pca.tex
pca tex finished. Improvements to pca in mesh.py

@@
 \documentclass[12pt]{article}
+\input{notation}
+\input{includes}
 
 \begin{document}
+\section{Principal Component Analysis}
+Given a collection of measured data points in a high dimensional space, it is often desirable to reduce the dimensionality of such data to a lower dimensional latent space, this could be to aid to processing the data or enable visualisation.
+In such instances, it is desirable for the mapping to the latent space to maintain as much information from the original data as possible.
 
-\section{Blendshapes and PCA}
+Principal Component Analysis (PCA) aims to reduce the dimensionality of a collection of data points while maximising the variance in the latent space.
+Alternatively, PCA aims to find a mapping from the original data to a new latent space which allow for the original data to be reconstructed with minimal error.
+These two aims are in fact equivalent.
 
+\subsection{Title?}
+Let $\bm{x_i} \in \mathbb{R}^f$ represent a data point in $f$ dimensional space and $\bm{y_i} \in \mathbb{R}^d$ represent the point which $\bm{x_i}$ is transformed to by a linear transformation $\mat{W}$ equation (\ref{eq:pca_transformation}), where $d << f$.
+
+\begin{equation} \label{eq:pca_transformation}
+    \bm{y_i} = \mat{W}^\top \bm{x_i}
+\end{equation}
+\noindent
+where,
+
+\begin{equation*}
+    \bm{y_i} = \begin{bmatrix} 
+                y_{i1} \\
+                \vdots \\
+                y_{id} 
+               \end{bmatrix},
+    \quad
+    \bm{x_i} = \begin{bmatrix} 
+                \bm{x}_{i1} \\
+                \vdots \\
+                \bm{x}_{if} 
+               \end{bmatrix},
+    \quad
+    \mat{W} = [
+               \bm{w_1}, \dots, \bm{w_d}
+              ],
+    \quad
+    \bm{w_i} = \begin{bmatrix} 
+                w_{i1} \\
+                \vdots \\
+                w_{if} 
+               \end{bmatrix}
+\end{equation*}
+
+\bigskip
+\noindent
+The optimal transformation $\mat{W}$ will maximise the variance in the data, where variance is expressed by equation (\ref{eq:variance}).
+This then follows that the optimal solution is found by maximising equation (\ref{eq:pca_optimum_transformation}).
+To prevent the trivial solution where $\bm{w}_k = \infty$, constrain a fixed magnitude $\mat{W}^\top \mat{W} = \mat{I}$.
+
+\begin{equation} \label{eq:variance}
+    \sigma_y^2 = \frac{1}{N} \sum_{i=1}^{N} (y_{ik} - \mu_k)^2
+\end{equation}
+\noindent
+where,
+\begin{equation*}
+    \mu_k = \frac{1}{N} \sum_{i=1}^N y_{ik}
+\end{equation*}
+
+\begin{align*}
+    \mat{W} &=\operatorname*{arg}\operatorname*{max}_\mat{W} 
+                \frac{1}{N} \sum_{k=1}^{d} \sum_{i=1}^{N} 
+                (y_{ik} - \mu_k)^2 \\
+            &=\operatorname*{arg}\operatorname*{max}_\mat{W} 
+                \frac{1}{N} \sum_{k=1}^{d} \sum_{i=1}^{N} 
+                \bm{w}_k^\top 
+                (\bm{x}_i - \bm{\mu})
+                (\bm{x}_i - \bm{\mu})^\top 
+                \bm{w}_k \\
+            &=\operatorname*{arg}\operatorname*{max}_\mat{W} 
+                \sum_{k=1}^{d}
+                \bm{w}_k^\top 
+                \mat{S}_t
+                \bm{w}_k \\
+\end{align*}
+\noindent
+where $\mat{S}_t$ is the covariance matrix,
+\begin{equation} \label{eq:covariance}
+    \mat{S}_t = \frac{1}{N} \sum_{i=1}^{N} 
+                (\bm{x}_i - \bm{\mu})
+                (\bm{x}_i - \bm{\mu})^\top 
+\end{equation}
+\noindent
+and,
+\begin{equation*}
+    \bm{\mu} = \frac{1}{N} \sum_{i=1}^N \bm{x}_i
+\end{equation*}
+
+\begin{equation} \label{eq:pca_optimum_transformation}
+    \mat{W} = \operatorname*{arg}\operatorname*{max}_\mat{W} 
+          \tr[\mat{W}^\top \mat{S}_t \mat{W}],
+          \quad
+    \text{subject to} \quad 
+    \mat{W}^\top \mat{W} = \mat{I}
+\end{equation}
+
+\bigskip
+\noindent
+By constructing the Lagangian from equation (\ref{eq:pca_optimum_transformation}) and solving the solution in equation (\ref{eq:pca_solved}) can be obtained.
+From eigendecomposition, $\mat{W}$ has columns of columns of eigenvectors which correspond to the $d$ largest non-zero eigenvalues of the covariance matrix, $\mat{S}_t$.
+
+\begin{equation*}
+    L(\mat{W}, \mat{\Lambda}) = \tr[\mat{W}^\top \mat{S}_t \mat{W}] - \tr[\Lambda(\mat{W}^\top \mat{W} - \mat{I})]
+\end{equation*}
+
+\begin{equation*}
+    \frac{\partial \tr[\mat{W}^\top \mat{S}_t \mat{W}]}{\partial \mat{W}} = 2 \mat{S}_t \mat{W},
+    \quad
+    \frac{\partial \tr[\Lambda(\mat{W}^\top \mat{W} - \mat{I})]}{\partial \mat{W}} = 2 \mat{W \Lambda}
+\end{equation*}
+
+\begin{equation*}
+    L(\mat{W}, \mat{\Lambda}) = 0 
+\end{equation*}
+
+\begin{equation} \label{eq:pca_solved}
+    \mat{S}_t \mat{W} = \mat{W \Lambda}
+\end{equation}
+
+\newpage
+\bibliographystyle{unsrt}
+\bibliography{ref}
 \end{document}

### 2019-06-20 Report/procrustes.tex
pca tex finished. Improvements to pca in mesh.py

@@
-\documentclass[12pt]{article}
+%\documentclass[12pt]{article}
 % some macros
-\input{notation}
-\input{includes}
+%\input{notation}
+%\input{includes}
 
-\begin{document}
+%\begin{document}
 
@@
 Procrustes Analysis is a statistical shape analysis method used to analyse the differences between two objects \cite{Krzanowski2000}.
-Given matrices $\mat{X} \in \mathbb{R}^{(n\times f)}$ and $\mat{Y} \in \mathbb{R}^{(n\times f)}$, each representing the coordinates of $n$ points in an $f$-dimensional feature space, a common statistical difference metric is the sum of square differences \ref{eq:ssd}.
+Given matrices $\mat{X} \in \mathbb{R}^{(n\times f)}$ and $\mat{Y} \in \mathbb{R}^{(n\times f)}$, each representing the coordinates of $n$ points in an $f$-dimensional feature space, a common statistical difference metric is the sum of square differences (\ref{eq:ssd}).
 \begin{equation}
@@
 This section follows the steps of object alignment using Procrustes Analysis as described above with a simple example.
-A pair of configurations are given by the matrices below, visualised in figure \ref{fig:procrustes_ex_1}.
+A pair of configurations are given by the matrices below, visualised in Figure \ref{fig:procrustes_ex_1}.
 
@@
 
-\newpage
-\bibliographystyle{unsrt}
-\bibliography{ref}
-\end{document}
+%\newpage
+%\bibliographystyle{unsrt}
+%\bibliography{ref}
+%\end{document}

### 2019-06-20 src/mesh/mesh.py

@@ -100,10 +100,17 @@ class Mesh():
             save_path = os.path.join(filepath, filename + '%05d' % file)
             self.export_mesh(vertices[:,file], faces, filename=save_path)
     
-    def create_blendshapes(self, vertices, n_shapes):
+    def create_blendshapes(self, vertices, n_shapes=None):
         # perform PCA on vertices to obtrain blendshapes
         v = vertices.T @ vertices
         eigvals, eigvecs = la.eig(v)    # Eigen analysis
+        eigvecs = np.real(eigvecs)
+        eigvals = np.real(eigvals)
+
+        # Remove zero eigen values
+        zero_idx = np.where(eigvals < 10E-10)[0]
+        eigvecs = np.delete(eigvecs, zero_idx, axis=1)
+        eigvals = np.delete(eigvals, zero_idx)
 
         # Sort eigenvectors by eigenvalues
         idx = eigvals.argsort()[::-1]   
@@ -114,11 +121,11 @@ class Mesh():
         eigvals_diag = np.diag(eigvals)
 
         # Create the blendshapes transformation matrix
-        transform = vertices @ eigvecs @ la.inv(eigvals_diag)
-        transform = np.real(transform)
+        blendshapes = vertices @ eigvecs @ la.inv(eigvals_diag)
 
         # Reduce down to amount desired
-        blendshapes = np.delete(transform, np.s_[n_shapes:], axis=1)
+        if n_shapes:
+            blendshapes = np.delete(blendshapes, np.s_[n_shapes:], axis=1)
 
         return blendshapes
     
@@ -183,7 +190,7 @@ class Mesh():
     
     
 if __name__ == "__main__":
-
+    """
     # Import root mesh to align onto
     root_mesh_dir = os.path.join("/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01")
     root_mesh = Mesh(root_mesh_dir, given_mesh="sentence01.000001.ply")
@@ -211,7 +218,7 @@ if __name__ == "__main__":
             os.makedirs(save_path)
         file_name = 'aligned'
         mesh.export_all_meshes(mesh_vertices, save_path, file_name)
-
+    """
     # Load aligned meshes to create blendshapes
     dir_plys = 'sentence01'
     dir_path = os.path.dirname(os.path.realpath(__file__))
@@ -222,8 +229,11 @@ if __name__ == "__main__":
 
[…]

### 2019-06-20 src/mesh/align_meshes.py
added a few scripts to manipulate meshes

@@ -1,0 +1,33 @@
+import os
+import numpy as np
+from mesh import Mesh
+
+if __name__ == "__main__":
+
+    # Import root mesh to align onto
+    root_mesh_dir = os.path.join("/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/sentence01")
+    root_mesh = Mesh(root_mesh_dir, given_mesh="sentence01.000001.ply")
+    root_mesh_vertices = root_mesh.get_empty_vertices(root_mesh.num_files)
+    root_mesh.get_vertex_postions(root_mesh_vertices)
+
+    # Align meshes
+    for file in range(1, 2):
+        sentence = 'sentence' + '%02d' %file
+
+        dir_plys = os.path.join("/home/peter/Documents/Uni/Project/datasets/registereddata/FaceTalk_170725_00137_TA/", sentence)
+        dir_path = os.path.dirname(os.path.realpath(__file__))
+
+        mesh = Mesh(dir_plys)
+        mesh_vertices = mesh.get_empty_vertices(mesh.num_files)
+        mesh.get_vertex_postions(mesh_vertices)
+
+        mesh.mesh_alignment(mesh_vertices, root_mesh_vertices)
+
+        mesh_vertices = mesh.vertices_to_2d(mesh_vertices)
+
+        # Export aligned meshes
+        save_path = os.path.join(dir_path, sentence)
+        if not os.path.exists(save_path):
+            os.makedirs(save_path)
+        file_name = 'aligned'
+        mesh.export_all_meshes(mesh_vertices, save_path, file_name)

### 2019-06-20 src/mesh/blendshape_interpolate.py

@@ -13,9 +13,12 @@ mesh_vertices = mesh.vertices_to_2d(mesh_vertices)
 verts = mesh_vertices[:,0]
 
 shapes = np.loadtxt(os.path.join(dir_path, 'shapes00.txt'), delimiter=',')
-shapes = normalize(shapes, axis=1)
+total_shapes = shapes.shape[1]
+n_shapes = 10
+shapes = np.delete(shapes, range(10, total_shapes), axis=1)
+#shapes = normalize(shapes, axis=1)
 
-save_path = os.path.join(dir_path, 'blendshapes/shape01')
+save_path = os.path.join(dir_path, 'blendshapes/shape00')
 if not os.path.exists(save_path):
     os.makedirs(save_path)
 

### 2019-06-20 src/mesh/create_blendshapes.py

@@ -1,0 +1,19 @@
+import os
+import numpy as np
+from mesh import Mesh
+
+if __name__ == "__main__":
+    # Load aligned meshes to create blendshapes
+    dir_plys = 'sentence01'
+    dir_path = os.path.dirname(os.path.realpath(__file__))
+    mesh = Mesh(os.path.join(dir_path, dir_plys))
+    mesh_vertices = mesh.get_empty_vertices(mesh.num_files)
+    mesh.get_vertex_postions(mesh_vertices)
+    mesh_vertices = mesh.vertices_to_2d(mesh_vertices)
+
+    frame_deltas = mesh.create_frame_deltas(mesh_vertices)
+
+    # vertices should have shape (features x samples)
+    shapes = mesh.create_blendshapes(frame_deltas)
+    print(shapes.shape)
+    np.savetxt('shapes00.txt', shapes, delimiter=',')

### 2019-06-20 src/mesh/recover_blendshape_params.py

@@ -1,0 +1,26 @@
+import os
+import numpy as np
+from mesh import Mesh
+
+if __name__ == "__main__":
+    # Load mesh to interpolate
+    dir_plys = 'sentence01'
+    dir_path = os.path.dirname(os.path.realpath(__file__))
+    mesh = Mesh(os.path.join(dir_path, dir_plys))
+    mesh_vertices = mesh.get_empty_vertices(mesh.num_files)
+    mesh.get_vertex_postions(mesh_vertices)
+    mesh_vertices = mesh.vertices_to_2d(mesh_vertices)
+    verts = mesh_vertices[:,0]
+
+    shapes = np.loadtxt(os.path.join(dir_path, 'shapes00.txt'), delimiter=',')
+    total_shapes = shapes.shape[1]
+    n_shapes = 5
+    shapes = np.delete(shapes, range(n_shapes, total_shapes), axis=1)
+
+    # Define blendshape parameters to interpolate along blendshpe axis.
+    blendshape_params = np.random.rand(5)
+    morphed_verts = mesh.morph_mesh(verts, shapes, blendshape_params)
+
+    params = mesh.recover_blendshape_parameters(morphed_verts, shapes)
+    print(params)
+    print(blendshape_params)

### 2019-06-21 src/audio/melspec.py
able to export a sequence of blendshapes for a given mesh sequence

@@ -1,0 +1,96 @@
+import numpy as np
+import librosa
+import librosa.display
+import matplotlib.pyplot as plt
+import os
+from scipy.io import wavfile
+from tqdm import tqdm
+
+class MelSpec():
+    def __init__(self, file_path=None, save_path=None):
+        self.file_path = file_path
+        self.save_path = save_path
+
+        if self.save_path and not os.path.exists(self.save_path):
+            os.makedirs(self.save_path)
+        
+        self.wav_files = []
+        for root, _dirs, files in os.walk(self.file_path):
+            for file in files:
+                if file.endswith('.wav'):
+                    self.wav_files.append(os.path.join(root, file))
+        self.wav_files = sorted(self.wav_files)
+        self.n_files = len(self.wav_files)
+
+    def read_wav(self, wav_file):
+        """
+        Reads a wav file and normalised based on the maximum amplitude.
+        """
+        if not wav_file.endswith('.wav'):
+            wav_file += '.wav'
+        wav_file = os.path.join(self.file_path, wav_file)
+
+        sample_rate, audio_data = wavfile.read(wav_file)
+        audio_data = audio_data / audio_data.max()
+
+        return sample_rate, audio_data
+    
+    def display_waveplot(self, audio, sample_rate):
+        """
+        Displays waveplot of audiofile
+        """
+        librosa.display.waveplot(y=audio, sr=sample_rate)
+        plt.show()
+    
+    def display_melplot(self, mel_spec):
+        """
+        Display mel spectogram
+        """
+        librosa.display.specshow(librosa.power_to_db(mel_spec), x_axis='time')
+        plt.show()
+    
+    def mfcc(self, audio_data, sample_rate, n_filters=30):
+        """
+        Returns the mel spectrum of an audio signal.
+        The number of mel filters can be varied.
+        """
+        mel_spec = librosa.feature.mfcc(y=audio_data, sr=sample_rate, n_mfcc=50)
+        return mel_spec
+
+    def save_spectrum(self, spectrum, filename):
+        """
+        Exports spectrum file
+        """
+        save_path = os.path.join(self.save_path, filename)
+        np.save(save_path, spectrum)
+    
+    def convert_to_mfcc(self):
+        """
+        Converts all wav files into mel spectograms and exports them as numpy 
+        array files.
+        Files are saved to self.save_path
[…]

### 2019-06-21 src/mesh/mesh.py

@@ -2,10 +2,15 @@ from plyfile import PlyData, PlyElement
 import scipy.linalg as la
 import os
 import numpy as np
+from tqdm import tqdm
 
 class Mesh():
-    def __init__(self, dir_plys, given_mesh=None):
+    def __init__(self, dir_plys, save_path=None, given_mesh=None):
         self.dir_plys = dir_plys
+        self.save_path = save_path
+
+        if self.save_path and not os.path.exists(self.save_path):
+            os.makedirs(self.save_path)
 
         if given_mesh:
             self.ply_files = [os.path.join(dir_plys, given_mesh)]
@@ -19,6 +24,11 @@ class Mesh():
         self.ply_files = sorted(self.ply_files)
         self.num_files = len(self.ply_files)
         self._get_mesh_metadata(self.ply_files[0])
+        self._get_root_mesh(self.ply_files[0])
+
+    def _get_root_mesh(self, root_ply):
+        self.root_mesh = self.get_empty_vertices(1)
+        self.get_vertex_postions(self.root_mesh)
 
     def _get_mesh_metadata(self, example_ply):
         with open(example_ply, 'rb') as f:
@@ -42,14 +52,16 @@ class Mesh():
         """
         Extracts vertex postions for all ply files
         """
-        print("Number of files being processed: {}".format(self.num_files))
+#        print("Number of files being processed: {}".format(self.num_files))
 
-        for file_number, file in enumerate(self.ply_files):
+        for file_number, file in tqdm(enumerate(self.ply_files)):
+            if file_number == target_array.shape[2]:
+                break
             with open(file, 'rb') as f:
                 plydata = PlyData.read(f)
                 self._get_vertex_postions(plydata, target_array, file_number)
-                if (file_number % 10) == 0:
-                    print("Processing file number: {}".format(file_number))
+                #if (file_number % 10) == 0:
+                #    print("Processing file number: {}".format(file_number))
     
     def vertices_to_2d(self, vertices):
         """
@@ -59,7 +71,7 @@ class Mesh():
         output shape (3*n_vertices, n_files)
         """
 
-        return vertices.reshape(3*self.vertex_count, self.num_files)
+        return vertices.reshape(3*self.vertex_count, -1)
     
[…]

### 2019-06-21 src/audio/audio_crop.sh
bash script to slice wav files into 0.1s clips

@@ -1,0 +1,59 @@
+#!/bin/bash
+# This script takes wav files under the SAMPLES directory and trims them
+#   into 0.1 second clips, the results are exported to OUTPUT
+
+# Source file path
+DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" >/dev/null 2>&1 && pwd )"
+
+# Source of original audio samples
+SAMPLES="$DIR/samples"
+
+# Trimmed output path
+OUTPUT="$DIR/trimmed_samples"
+
+# File extension
+EXTENSION=".wav"
+
+# Crop duration
+DURATION=0.1
+
+# Temp file name
+TEMP_FILE="temp_file"$EXTENSION
+
+# Make filepath for trimmed audio samples to live
+mkdir -p $OUTPUT
+
+for file in $SAMPLES/*.wav
+do
+    filename="${file%.*}"
+    filename="${filename##*/}"
+
+    # Backup file
+    backup=$SAMPLES/$filename"_bkup"$EXTENSION
+    cp $file $backup
+
+    # Get duration of file
+    file_length="$(soxi -D $file)"
+
+    output_path=$OUTPUT/$filename
+    mkdir -p $output_path
+
+    cp $file $TEMP_FILE
+
+    # Get number of slices the audio clip will be cut into.
+    let "slices = $(bc <<< "$file_length / $DURATION")"
+
+    for i in $(seq -f "%04g" 1 $slices)
+    do
+        sox $file $output_path/$filename_$i$EXTENSION trim 0 $DURATION
+        sox $file $TEMP_FILE trim 0.1
+        mv $TEMP_FILE $file
+    done
+
+    # Recover original from backup
+    rm $file
+    mv $backup $file
+done
+
+# File can be reconstructed by calling from the dir containing the sentence
+#   sox *.wav out.wav

### 2019-06-21 src/audio/melspec.py

@@ -83,14 +83,15 @@ if __name__ == "__main__":
     file_path = 'samples'
     spectogram_path = 'spectograms'
     file_name = 'sentence01'
+    file_name = 'temp'
 
     audio_files = os.path.join(dir_path, file_path)
     spec_files = os.path.join(dir_path, spectogram_path)
 
-    ms = MelSpec(file_path=audio_files, save_path=spec_files)
+    ms = MelSpec(file_path=dir_path, save_path=dir_path)
     #ms.convert_to_mfcc()
     sr, audio_data = ms.read_wav(file_name)
     mel_spec = ms.mfcc(audio_data, sr)
     print(mel_spec.shape)
-    #ms.display_melplot(mel_spec)
+    ms.display_melplot(mel_spec)
     #ms.save_spectrum(mel_spec, file_name)

### 2019-06-21 src/audio/audio_crop.sh
check for edge cases where there is no audio

@@ -20,6 +20,8 @@ DURATION=0.1
 # Temp file name
 TEMP_FILE="temp_file"$EXTENSION
 
+THRESHOLD=0.00000001
+
 # Make filepath for trimmed audio samples to live
 mkdir -p $OUTPUT
 
@@ -45,9 +47,18 @@ do
 
     for i in $(seq -f "%04g" 1 $slices)
     do
-        sox $file $output_path/$filename_$i$EXTENSION trim 0 $DURATION
+        outfile=$output_path/$filename_$i$EXTENSION
+        sox $file $outfile trim 0 $DURATION
         sox $file $TEMP_FILE trim 0.1
         mv $TEMP_FILE $file
+
+        # Check that the amplitude of the file isn't zero
+        amplitude="$(sox $outfile -n stat 2>&1 | sed -n 's#^Maximum amplitude:[^0-9]*\([0-9.]*\)$#\1#p')"
+        if (( $(echo "$amplitude < $THRESHOLD" |bc -l) )); then
+            echo $outfile
+            rm $outfile
+            break
+        fi
     done
 
     # Recover original from backup

### 2019-07-29 src/mesh/recover_blendshape_params.py
able to train shape lrw classifer

@@ -4,7 +4,6 @@ import argparse
 from mesh import Mesh
 from mesh import gen_file_list
 from tqdm import tqdm
-import shutil
 
 if __name__ == "__main__":
 
@@ -16,7 +15,7 @@ if __name__ == "__main__":
     parser.add_argument('--params_name', default='shape_params')
     parser.add_argument('--root_mesh_dir', 
                         default='root_meshes/voca_root/')
-    parser.add_argument('--recover_from_dir', default='lrw_subset_meshes/ABOUT/train/ABOUT_00001')
+    parser.add_argument('--recover_from_dir', default='lrw_subset_meshes/')
 
     args = parser.parse_args()
 
@@ -36,6 +35,7 @@ if __name__ == "__main__":
     root_mesh_list = gen_file_list(root_mesh_path, ext='.ply')
     root_mesh = Mesh(root_mesh_list)
 
+    """
     dir_plys = args.recover_from_dir
     dir_plys_path = os.path.join(dir_path, dir_plys)
     file_name = dir_plys.split('/')[-2]
@@ -68,20 +68,18 @@ if __name__ == "__main__":
                 mesh_vertices[:,v], shapes)
 
         np.save(file_save_path, params)
-        #shutil.rmtree(dir_plys_path)
     else:
         print("File exists")
-
-
-
-
     """
+
+
+
     dir_plys = args.recover_from_dir
     all_subjects_path = os.path.join(dir_path, dir_plys)
     subjects_list = sorted(os.listdir(all_subjects_path))
 
     for subject in tqdm(subjects_list):
-        subject_path = os.path.join(all_subjects_path, subject)
+        subject_path = os.path.join(all_subjects_path, subject, 'train')
         sentence_list = sorted(os.listdir(subject_path))
         for sentence_num, sentence in enumerate(sentence_list):
             ply_files = os.path.join(subject_path, sentence)
@@ -113,4 +111,3 @@ if __name__ == "__main__":
                     mesh_vertices[:,v], shapes)
 
             np.save(file_save_path, params)
-    """

### 2019-07-29 src/model/base/base_data_loader.py

@@ -1,6 +1,6 @@
 import numpy as np
 from torch.utils.data import DataLoader
-from torch.utils.data.sampler import SequentialSampler
+from torch.utils.data.sampler import SequentialSampler, RandomSampler
 
 class BaseDataLoader(DataLoader):
     def __init__(self, dataset, batch_size, shuffle, 
@@ -29,14 +29,13 @@ class BaseDataLoader(DataLoader):
 
         all_idx = np.arange(self.n_samples)
         if self.shuffle:
-            np.random.seed(0)
             np.random.shuffle(all_idx)
         
         val_idx = all_idx[0:self.n_val]
         train_idx = np.delete(all_idx, np.arange(0, self.n_val))
 
-        train_sampler = SequentialSampler(train_idx)
-        val_sampler = SequentialSampler(val_idx)
+        train_sampler = RandomSampler(train_idx)
+        val_sampler = RandomSampler(val_idx)
 
         self.shuffle = False
 

### 2019-07-29 src/model/base/base_trainer.py

@@ -66,3 +66,62 @@ class BaseMultiTrainer(BaseDiscriminator, BaseGenerator):
     @abstractmethod
     def train(self):
         raise NotImplementedError
+
+class BaseTrainer():
+    def __init__(self, config, model, loss_func, optimizer):
+        self.config = config
+        self.logger = config.get_logger('trainer', config['trainer']['level'])
+        self.device, device_ids = self._prepare_device(config['n_gpu'])
+
+        self.model = model.to(self.device)
+        if len(device_ids) > 1:
+            self.model = torch.nn.DataParallel(model, device_ids=device_ids)
+
+        self.loss_func = loss_func
+        self.optimizer = optimizer
+        self.epochs = config['trainer']['epochs']
+
+        self.save_period = config['trainer']['save_period']
+        self.checkpoint_dir = config.save_dir
+        # setup visualization writer instance                
+        self.writer = TensorboardWriter(config.log_dir, self.logger, 
+                                        config['trainer']['tensorboard'])
+    
+    @abstractmethod
+    def _training_epoch(self, epoch):
+        raise NotImplementedError
+
+    @abstractmethod
+    def train(self):
+        raise NotImplementedError
+    
+    def _save_checkpoint(self, epoch):
+        arch = type(self.model).__name__
+        state = {
+            'arch': arch,
+            'epoch': epoch,
+            'state_dict': self.model.state_dict(),
+            'optimizer': self.optimizer.state_dict(),
+            'config': self.config
+        }
+        filename = str(self.checkpoint_dir / 'checkpoint-{}.pth'.format(epoch))
+        torch.save(state, filename)
+
+    def _resume_checkpoint(self, resume_path):
+        resume_path = str(resume_path)
+        checkpoint = torch.load(resume_path)
+        self.start_epoch = checkpoint['epoch'] + 1
+
+        self.model.load_state_dict(checkpoint['state_dict'])
+
+        self.optimizer.load_state_dict(checkpoint['optimizer'])
+
+    def _prepare_device(self, n_gpu_use):
+        n_gpu = torch.cuda.device_count()
+        if n_gpu_use > 0 and n_gpu == 0:
+            n_gpu_use = 0
+        if n_gpu_use > n_gpu:
+            n_gpu_use = n_gpu
+        device = torch.device('cuda:0' if n_gpu_use > 0 else 'cpu')
+        list_ids = list(range(n_gpu_use))
+        return device, list_ids 

### 2019-07-29 src/model/data_loader/data_loaders.py

@@ -86,3 +86,32 @@ class DataLoaderWavShapes(BaseDataLoader):
 
         super().__init__(self.dataset, batch_size, shuffle,
                          train_split, n_workers, drop_last=drop_last)
+
+class DataLoaderLrwShapes(BaseDataLoader):
+    """
+    LRW Shape parameters DataLoader
+    """
+    def __init__(self, blendshapes_dir, batch_size, shuffle=True,
+                 train_split=1.0, n_workers=0, drop_last=True, 
+                 tsfm=None):
+                
+        self.blendshapes_dir = blendshapes_dir
+        self.batch_size = batch_size
+        self.shuffle = shuffle
+        self.train_split = train_split
+        self.n_workers = n_workers
+        self.drop_last = drop_last
+
+        transform_list = []
+        for t in tsfm:
+            item = getattr(datasets, t)()
+            transform_list.append(item)
+        
+        transform = transforms.Compose(transform_list)
+
+        self.dataset = LrwBlendshapesDataset(self.blendshapes_dir,
+                                             transform=transform)
+
+        super().__init__(self.dataset, self.batch_size, self.shuffle,
+                         self.train_split, self.n_workers, 
+                         drop_last=self.drop_last)

### 2019-07-29 src/model/datasets/datasets.py

@@ -1,4 +1,5 @@
 import torch
+import torch.tensor as tensor
 from torch.utils.data import Dataset
 import os
 import math
@@ -7,6 +8,89 @@ from scipy.io import wavfile
 import librosa
 import numpy as np
 from tqdm import tqdm
+
+
+class LrwBlendshapesDataset(Dataset):
+    """
+    Blendshape Parameters for audio clips from the BBC LRW words dataset,
+    processed by the VOCA model.
+
+    500 words (labels), 43 frames for 1 second, 4 shape parameters per frame.
+    """
+    def __init__(self, blendshapes_path, transform=None):
+        self.blendshapes_path = blendshapes_path
+        self.blendshapes_list = sorted(os.listdir(self.blendshapes_path))
+        self.transform = transform
+
+        self.stats = {'shape_min': np.inf,
+                      'shape_max': -np.inf}
+
+        self._collect_stats()
+
+    def __len__(self):
+        return len(self.blendshapes_list)
+    
+    def __getitem__(self, idx):
+        """
+        idx references the sample number
+        """
+        shape_params, label = self._get_sample(idx)
+
+        sample = {'shape_params': shape_params.astype(np.float32),
+                  'label': label.astype(np.int64)}
+
+        sample = self._normalize(sample)
+
+        if self.transform:
+            sample = self.transform(sample)
+        
+        return sample
+    
+    def _get_sample(self, idx):
+        """
+        fetches the relevant sample
+        """
+        sample_file = self.blendshapes_list[idx]
+        sample_name = sample_file.split('.')[0]
+        word_label, word_id = sample_name.split('_')
+        word_label = np.array(int(word_label))
+        shape_params = np.load(os.path.join(self.blendshapes_path, sample_file))
+        return shape_params, word_label
+
+
+    def _collect_stats(self):
+        """
+        Finds max and min values for blendshape params for dataset.
+        """
+        print("Collecting dataset statistics...\n")
+        for idx in tqdm(range(self.__len__())):
+            shape_param, _label = self._get_sample(idx)
+
+            idx_shape_min = shape_param.min()
+            idx_shape_max = shape_param.max()
+
+            if idx_shape_min < self.stats['shape_min']: 
+                self.stats['shape_min'] = idx_shape_min
+            if idx_shape_max > self.stats['shape_max']: 
+                self.stats['shape_max'] = idx_shape_max
+
[…]

### 2019-07-29 src/model/model/models.py

@@ -445,3 +445,59 @@ class MFCC_Shape_Critic(BaseModel):
         x = self.lrelu8(self.conv8(x))
         x = self.lin9(self.conv9(x))
         return x
+
+class Lrw_Shape_Classifier(BaseModel):
+    def __init__(self):
+        super().__init__()
+
+        self.conv1 = nn.Conv2d(1, 16, kernel_size=2)
+        self.relu1 = nn.ReLU()
+
+        self.conv2 = nn.Conv2d(16, 32, kernel_size=2)
+        self.relu2 = nn.ReLU()
+
+        self.conv3 = nn.Conv2d(32, 64, kernel_size=2)
+        self.relu3 = nn.ReLU()
+
+        self.conv4 = nn.Conv1d(64, 128, kernel_size=6)
+        self.relu4 = nn.ReLU()
+
+        self.conv5 = nn.Conv1d(128, 256, kernel_size=6)
+        self.relu5 = nn.ReLU()
+
+        self.conv6 = nn.Conv1d(256, 256, kernel_size=6)
+        self.relu6 = nn.ReLU()
+
+        self.conv7 = nn.Conv1d(256, 256, kernel_size=6)
+        self.relu7 = nn.ReLU()
+
+        self.conv8 = nn.Conv1d(256, 256, kernel_size=6)
+        self.relu8 = nn.ReLU()
+
+        self.conv9 = nn.Conv1d(256, 256, kernel_size=6)
+        self.relu9 = nn.ReLU()
+
+        self.linear10 = nn.Linear(2560, 500)
+        self.softmax10 = nn.functional.softmax
+
+
+    
+    def forward(self, shapes):
+        batch_size = shapes.size(0)
+
+        x = self.relu1(self.conv1(shapes))
+        x = self.relu2(self.conv2(x))
+        x = self.relu3(self.conv3(x))
+
+        x = x.squeeze(2)
+        x = self.relu4(self.conv4(x))
+        x = self.relu5(self.conv5(x))
+        x = self.relu6(self.conv6(x))
+        x = self.relu7(self.conv7(x))
+        x = self.relu8(self.conv8(x))
+        x = self.relu9(self.conv9(x))
+
+        x = x.view(batch_size, -1)
+        x = self.softmax10(self.linear10(x), dim=1)
+
+        return x

### 2019-07-29 src/model/train.py

@@ -1,7 +1,7 @@
 import torch
 from model import losses, models
 from data_loader import data_loaders
-from trainer import MFCCShapeTrainer, CGANTrainer
+from trainer import MFCCShapeTrainer, LrwShapeTrainer
 from get_config import GetConfig
 from utils import fix_seed
 import torch.optim as optim
@@ -39,9 +39,32 @@ def gan_main(config):
                          gen_model, gen_loss, gen_optimizer)
     trainer.train()
 
+def classify_main(config):
+    logger = config.get_logger('train')
+
+    train_loader = config.get('data_loader', data_loaders)
+    val_loader = train_loader.val_split()
+
+    model = config.get('arch', models)
+
+    logger.info(model)
+
+    trainable_params = filter(lambda p: p.requires_grad, model.parameters())
+
+    optimizer = config.get('optimizer', optim, trainable_params)
+
+    loss = config.get_func('loss_func', losses)
+
+    #TODO add Trainer selection to config
+    trainer = LrwShapeTrainer(config, train_loader, model,
+                              loss, optimizer, val_loader)
+    trainer.train()
+
 if __name__ == "__main__":
     fix_seed(0)
 
-    config = GetConfig('./config/mfcc_shape_gan/config.json')
+    #config = GetConfig('./config/mfcc_shape_gan/config.json')
+    #gan_main(config)
 
-    gan_main(config)
+    config = GetConfig('./config/lrw_shape_classifier/config.json')
+    classify_main(config)

### 2019-07-29 src/model/trainer/trainer.py

@@ -1,6 +1,6 @@
 import torch
 import numpy as np
-from base import BaseMultiTrainer
+from base import BaseMultiTrainer, BaseTrainer
 from logger import TensorboardWriter
 from torchvision.utils import save_image
 from model.losses import gradient_penalty
@@ -345,7 +345,7 @@ class MFCCShapeTrainer(BaseMultiTrainer):
 
         self.data_loader = data_loader
         self.vis_loader = vis_loader
-        self.batch_size = data_loader.batch_size
+        self.batch_size = self.data_loader.batch_size
         self.vis_batch_size = vis_loader.batch_size
         self.len_epoch = len(self.data_loader)
 
@@ -466,3 +466,115 @@ class MFCCShapeTrainer(BaseMultiTrainer):
         sample_name = 'generated_sample_epoch_%03d' % epoch
         save_dir = self.config.samples_dir / sample_name
         np.save(save_dir, gen_sample)
+
+class LrwShapeTrainer(BaseTrainer):
+    def __init__(self, config, train_loader,
+                 model, loss_func, optimizer, val_loader=None):
+    
+        self.train_loader = train_loader
+        self.val_loader = val_loader
+
+        self.batch_size = self.train_loader.batch_size
+
+        self.len_train_epoch = len(self.train_loader)
+
+        if val_loader:
+            self.val_step = True
+            self.len_val_epoch = len(self.val_loader)
+        else:
+            self.val_step = False
+
+        self.log_step = int(np.sqrt(self.batch_size))
+
+        super().__init__(config, model, loss_func, optimizer)
+    
+    def train(self):
+        for epoch in range(1, self.epochs+1):
+            log = self._training_epoch(epoch)
+
+            print(log['train_loss'])
+            print(log['val_loss'])
+
+            if epoch % self.save_period == 0:
+                self._save_checkpoint(epoch)
+    
+    def _training_epoch(self, epoch):
+        self.model.train()
+
+        total_loss = 0
+        correct = 0
+
+        for batch_idx, sample in enumerate(self.train_loader):
+            step = (epoch-1) * self.len_train_epoch + batch_idx
+
+            label = sample['label']
+            label = label.to(self.device)
+
+            shape_params = sample['shape_params']
+            shape_params = shape_params.to(self.device)
+
+            self.optimizer.zero_grad()
+            output = self.model(shape_params)
+            preds = output.argmax(dim=1, keepdim=True)
+            correct += preds.eq(label.view_as(preds)).sum().item()
[…]

### 2019-07-30 src/model/model/models.py
lrw classifier. Doesn't work

@@ -477,7 +477,7 @@ class Lrw_Shape_Classifier(BaseModel):
         self.conv9 = nn.Conv1d(256, 256, kernel_size=6)
         self.relu9 = nn.ReLU()
 
-        self.linear10 = nn.Linear(2560, 500)
+        self.linear10 = nn.Linear(2560, 2)
         self.softmax10 = nn.functional.softmax
 
 

### 2019-07-30 src/model/trainer/trainer.py

@@ -492,9 +492,6 @@ class LrwShapeTrainer(BaseTrainer):
         for epoch in range(1, self.epochs+1):
             log = self._training_epoch(epoch)
 
-            print(log['train_loss'])
-            print(log['val_loss'])
-
             if epoch % self.save_period == 0:
                 self._save_checkpoint(epoch)
     
@@ -503,6 +500,7 @@ class LrwShapeTrainer(BaseTrainer):
 
         total_loss = 0
         correct = 0
+        val_loss = 0
 
         for batch_idx, sample in enumerate(self.train_loader):
             step = (epoch-1) * self.len_train_epoch + batch_idx
@@ -538,8 +536,6 @@ class LrwShapeTrainer(BaseTrainer):
                     'Accuracy: {:.6f}'.format(epoch, batch_idx, loss, acc))
 
         train_loss = total_loss / self.len_train_epoch
-        train_accuracy = correct / self.len_train_epoch
-        print("Train Accuracy: {:.4f}".format(train_accuracy))
 
         if self.val_step:
             val_loss = self._val_epoch(epoch)
@@ -573,8 +569,5 @@ class LrwShapeTrainer(BaseTrainer):
                           'Loss: {:.6f} Accuracy: {:.6f}'.format(
                               epoch, batch_idx, loss, acc))
 
-
         val_loss = total_loss / self.len_val_epoch
-        val_accuracy = correct / self.len_val_epoch
-        print("Validation Accuracy: {:.4f}".format(val_accuracy))
         return val_loss

### 2019-07-30 src/model/base/base_trainer.py
lrw audio to blendshape gan ready to train

@@ -53,6 +53,29 @@ class BaseMultiTrainer(BaseDiscriminator, BaseGenerator):
             self.disc_model = torch.nn.DataParallel(self.disc_model, 
                                                     device_ids=device_ids)
 
+    def _save_checkpoint(self, epoch):
+        disc_arch = type(self.disc_model).__name__
+        disc_state = {
+            'arch': disc_arch,
+            'epoch': epoch,
+            'state_dict': self.disc_model.state_dict(),
+            'optimizer': self.disc_optimizer.state_dict(),
+            'config': self.config
+        }
+        disc_filename = f'{self.checkpoint_dir}/disc_checkpoint-{epoch}.pth'
+        torch.save(disc_state, disc_filename)
+
+        gen_arch = type(self.gen_model).__name__
+        gen_state = {
+            'arch': gen_arch,
+            'epoch': epoch,
+            'state_dict': self.gen_model.state_dict(),
+            'optimizer': self.gen_optimizer.state_dict(),
+            'config': self.config
+        }
+        gen_filename = f'{self.checkpoint_dir}/gen_checkpoint-{epoch}.pth'
+        torch.save(gen_state, gen_filename)
+
     def _prepare_device(self, n_gpu_use):
         n_gpu = torch.cuda.device_count()
         if n_gpu_use > 0 and n_gpu == 0:

### 2019-07-30 src/model/datasets/datasets.py

@@ -8,6 +8,7 @@ from scipy.io import wavfile
 import librosa
 import numpy as np
 from tqdm import tqdm
+import pickle
 
 
 class LrwBlendshapesDataset(Dataset):
@@ -91,7 +92,6 @@ class LrwBlendshapesDataset(Dataset):
         return {'shape_params': shape_params,
                 'label': label}
     
-
 class WavBlendshapesDataset(Dataset):
     """
     Raw Wav files (variable length)
@@ -100,7 +100,7 @@ class WavBlendshapesDataset(Dataset):
     This is currently using all blendshape parameters
     """
 
-    def __init__(self, wav_path, blendshapes_path, n_shapes=10, transform=None):
+    def __init__(self, wav_path, blendshapes_path, n_shapes=4, transform=None):
         self.wav_path = wav_path
         self.blendshapes_path = blendshapes_path
         self.n_shapes = n_shapes
@@ -136,6 +136,10 @@ class WavBlendshapesDataset(Dataset):
         shape_name = item_name + '.npy'
 
         sr, audio_data = wavfile.read(os.path.join(self.wav_path, wav_name))
+
+        if len(audio_data.shape) > 1:
+            audio_data = np.delete(audio_data, 1, 1).reshape(-1,)
+
         audio_data = audio_data / audio_data.max() 
         mfcc = self._mfcc(audio_data, sr)
 
@@ -149,24 +153,31 @@ class WavBlendshapesDataset(Dataset):
         """
         Finds max and min values for blendshape params and mfcc for dataset.
         """
-        print("Collecting dataset statistics...\n")
-        for idx in tqdm(range(self.__len__())):
-            mfcc, shape_param = self._get_data_pair(idx)
+        try:
+            with open('data/lrw_audio_stats.pkl', 'rb') as f:
+                self.stats = pickle.load(f)
+        except:
+            print("Collecting dataset statistics...\n")
+            for idx in tqdm(range(self.__len__())):
+                mfcc, shape_param = self._get_data_pair(idx)
 
-            idx_mfcc_min = mfcc.min()
-            idx_mfcc_max = mfcc.max()
-            idx_shape_min = shape_param.min()
-            idx_shape_max = shape_param.max()
+                idx_mfcc_min = mfcc.min()
+                idx_mfcc_max = mfcc.max()
+                idx_shape_min = shape_param.min()
+                idx_shape_max = shape_param.max()
 
[…]

### 2019-07-30 src/model/train.py

@@ -12,7 +12,8 @@ def gan_main(config):
     logger = config.get_logger('train')
 
     data_loader = config.get('data_loader', data_loaders)
-    vis_loader = config.get('vis_loader', data_loaders)
+    vis_loader = data_loader.val_split()
+    #vis_loader = config.get('vis_loader', data_loaders)
 
     disc_model = config.get('discriminator,arch', models)
 
@@ -63,8 +64,8 @@ def classify_main(config):
 if __name__ == "__main__":
     fix_seed(0)
 
-    #config = GetConfig('./config/mfcc_shape_gan/config.json')
-    #gan_main(config)
+    config = GetConfig('./config/mfcc_shape_gan/config.json')
+    gan_main(config)
 
-    config = GetConfig('./config/lrw_shape_classifier/config.json')
-    classify_main(config)
+    #config = GetConfig('./config/lrw_shape_classifier/config.json')
+    #classify_main(config)

### 2019-07-30 src/model/trainer/trainer.py

@@ -453,8 +453,8 @@ class MFCCShapeTrainer(BaseMultiTrainer):
                              'Gen Loss: {}'.format(disc_loss, gen_loss))
 
             self.save_sample(fixed_noise, fixed_mfcc, epoch)
-            #if epoch % self.save_period == 0:
-            #    self._save_checkpoint(epoch)
+            if epoch % self.save_period == 0:
+                self._save_checkpoint(epoch)
     
     def save_sample(self, noise, mfcc, epoch):
         gen_sample = self.gen_model(noise, mfcc).detach().to('cpu')

### 2019-07-30 src/model/base/base_data_loader.py
updating visualisation dataloader

@@ -31,15 +31,20 @@ class BaseDataLoader(DataLoader):
         if self.shuffle:
             np.random.shuffle(all_idx)
         
-        val_idx = all_idx[0:self.n_val]
-        train_idx = np.delete(all_idx, np.arange(0, self.n_val))
+        if self.n_val > 0:
+            val_idx = all_idx[0:self.n_val]
+            train_idx = np.delete(all_idx, np.arange(0, self.n_val))
 
-        train_sampler = RandomSampler(train_idx)
-        val_sampler = RandomSampler(val_idx)
+            train_sampler = RandomSampler(train_idx)
+            val_sampler = RandomSampler(val_idx)
+            samplers = (train_sampler, val_sampler)
+        else:
+            sampler = RandomSampler(all_idx)
+            samplers = (sampler, None)
 
         self.shuffle = False
 
-        return train_sampler, val_sampler
+        return samplers
     
     def val_split(self):
         return DataLoader(dataset=self.dataset, batch_size=self.batch_size, 

### 2019-07-30 src/model/train.py

@@ -12,8 +12,8 @@ def gan_main(config):
     logger = config.get_logger('train')
 
     data_loader = config.get('data_loader', data_loaders)
-    vis_loader = data_loader.val_split()
-    #vis_loader = config.get('vis_loader', data_loaders)
+    #vis_loader = data_loader.val_split()
+    vis_loader = config.get('vis_loader', data_loaders)
 
     disc_model = config.get('discriminator,arch', models)
 

### 2019-07-30 src/model/trainer/trainer.py

@@ -95,7 +95,6 @@ class GANTrainer(BaseMultiTrainer):
                             'Gen Loss: {:.6f} '.format(
                                 epoch, batch_idx+1, gen_loss))
 
-                self.writer.set_step(step)
                 self.writer.add_scalar('disc_loss', disc_loss)
                 self.writer.add_scalar('gen_loss', gen_loss)
             
@@ -154,6 +153,7 @@ class CGANTrainer(BaseMultiTrainer):
             disc_loss = disc_real_loss + disc_fake_loss + gp
         else:
             disc_loss = disc_real_loss + disc_fake_loss
+
         self.disc_model.zero_grad()
         disc_loss.backward()
         self.disc_optimizer.step()
@@ -376,9 +376,14 @@ class MFCCShapeTrainer(BaseMultiTrainer):
                                                  real=shape_param,
                                                  fake=fake_shapes,
                                                  conditional=mfcc)
+            self.writer.add_scalar('critic/gradient_penalty', gp)
+
             disc_loss = disc_real_loss + disc_fake_loss + gp
         else:
             disc_loss = disc_real_loss + disc_fake_loss
+
+        self.writer.add_scalar('critic/real_loss', disc_real_loss)
+        self.writer.add_scalar('critic/fake_loss', disc_fake_loss)
 
         self.disc_model.zero_grad()
         disc_loss.backward()
@@ -416,6 +421,7 @@ class MFCCShapeTrainer(BaseMultiTrainer):
             total_gen_loss = 0
             for batch_idx, sample in enumerate(self.data_loader):
                 step = (epoch-1) * self.len_train_epoch + batch_idx
+                self.writer.set_step(step)
 
                 mfcc = sample['mfcc'].to(self.device)
                 shape_param = sample['shape_param'].to(self.device)
@@ -429,7 +435,7 @@ class MFCCShapeTrainer(BaseMultiTrainer):
                     self.logger.info(
                         'Train Epoch: {} '
                         'Batch: {} '
-                        'Disc Loss: {:.6f} '.format(
+                        'Critic Loss: {:.6f} '.format(
                             epoch, batch_idx+1, disc_loss))
 
                 # Train Generator
@@ -443,13 +449,12 @@ class MFCCShapeTrainer(BaseMultiTrainer):
                             'Gen Loss: {:.6f} '.format(
                                 epoch, batch_idx+1, gen_loss))
 
-                self.writer.set_step(step)
[…]

### 2019-07-31 src/model/model/models.py
removed unused models. Updated mfcc-shapes gan model

@@ -3,218 +3,6 @@ import torch.nn as nn
 import torch.nn.functional as F
 from base import BaseModel
 
-class Generator_MNIST(BaseModel):
-    def __init__(self, z_dim=100):
-        super().__init__()
-
-        self.convt1 = nn.ConvTranspose2d(z_dim, 256, kernel_size=4, bias=False)
-        self.bn1 = nn.BatchNorm2d(256)
-        self.relu1 = nn.ReLU()
-
-        self.convt2 = nn.ConvTranspose2d(256, 128, kernel_size=3, bias=False)
-        self.bn2 = nn.BatchNorm2d(128)
-        self.relu2 = nn.ReLU()
-
-        self.convt3 = nn.ConvTranspose2d(128, 64, kernel_size=3, bias=False)
-        self.bn3 = nn.BatchNorm2d(64)
-        self.relu3 = nn.ReLU()
-
-        self.convt4 = nn.ConvTranspose2d(64, 64, kernel_size=4, bias=False)
-        self.bn4 = nn.BatchNorm2d(64)
-        self.relu4 = nn.ReLU()
-
-        self.convt5 = nn.ConvTranspose2d(64, 32, kernel_size=4, stride=2, bias=False)
-        self.bn5 = nn.BatchNorm2d(32)
-        self.relu5 = nn.ReLU()
-
-        self.convt6 = nn.ConvTranspose2d(32, 16, kernel_size=3, bias=False)
-        self.bn6 = nn.BatchNorm2d(16)
-        self.relu6 = nn.ReLU()
-
-        self.convt7 = nn.ConvTranspose2d(16, 1, kernel_size=3, bias=False)
-        self.tanh7 = nn.Tanh()
-
-
-    def forward(self, x):
-        x = x.view(x.size(0), x.size(1), 1, 1)
-        x = self.relu1(self.bn1(self.convt1(x)))
-        x = self.relu2(self.bn2(self.convt2(x)))
-        x = self.relu3(self.bn3(self.convt3(x)))
-        x = self.relu4(self.bn4(self.convt4(x)))
-        x = self.relu5(self.bn5(self.convt5(x)))
-        x = self.relu6(self.bn6(self.convt6(x)))
-        out = self.tanh7(self.convt7(x))
-        return out
-
-class Discriminator_MNIST(BaseModel):
-    def __init__(self):
-        super().__init__()
-
[…]

### 2019-07-31 src/model/model/models.py
changed gen model

@@ -123,7 +123,7 @@ class Voca_Shape_Critic(BaseModel):
 
         return out
 
-class MFCC_Shape_Gen(BaseModel):
+class old_MFCC_Shape_Gen(BaseModel):
     def __init__(self, z_dim, shapes_dim):
         super().__init__()
 
@@ -177,6 +177,50 @@ class MFCC_Shape_Gen(BaseModel):
         x = self.relu8(self.conv8(x))
         x = self.relu9(self.conv9(x))
         x = self.tanh10(self.conv10(x))
+        return x
+
+class MFCC_Shape_Gen(BaseModel):
+    def __init__(self, z_dim, shapes_dim):
+        super().__init__()
+
+        in_dim = z_dim + 1
+
+        self.conv1 = nn.Conv2d(in_dim, 32, kernel_size=(4,5), stride=(2,1))
+        self.relu1 = nn.ReLU()
+
+        self.conv2 = nn.Conv2d(32, 64, kernel_size=(4,5), stride=(2,1))
+        self.relu2 = nn.ReLU()
+
+        self.conv3 = nn.Conv2d(64, 128, kernel_size=(3,5), stride=(1,1))
+        self.relu3 = nn.ReLU()
+
+        self.conv4 = nn.Conv2d(128, 128, kernel_size=(3,4), stride=(1,1))
+        self.relu4 = nn.ReLU()
+
+        self.conv5 = nn.Conv2d(128, 128, kernel_size=(3,4), stride=(1,1))
+        self.relu5 = nn.ReLU()
+
+        self.conv6 = nn.Conv2d(128, 64, kernel_size=(3,3), stride=(1,1))
+        self.relu6 = nn.ReLU()
+
+        self.conv7 = nn.Conv2d(64, 32, kernel_size=(3,3), stride=(1,1))
+        self.relu7 = nn.ReLU()
+
+        self.convt8 = nn.ConvTranspose1d(32, 4, kernel_size=3, stride=2)
+        self.tanh8 = nn.Tanh()
+
+
+    def forward(self, noise, mfcc):
+        x = torch.cat((noise, mfcc), dim=1)
+        x = self.relu1(self.conv1(x))
+        x = self.relu2(self.conv2(x))
+        x = self.relu3(self.conv3(x))
+        x = self.relu4(self.conv4(x))
+        x = self.relu5(self.conv5(x))
+        x = self.relu6(self.conv6(x))
+        x = self.relu7(self.conv7(x))
[…]

### 2019-07-31 src/model/trainer/trainer.py

@@ -1,3 +1,4 @@
+import os
 import torch
 import numpy as np
 from base import BaseMultiTrainer, BaseTrainer
@@ -462,15 +463,23 @@ class MFCCShapeTrainer(BaseMultiTrainer):
                 self._save_checkpoint(epoch)
     
     def save_sample(self, noise, mfcc, epoch):
+
         gen_sample = self.gen_model(noise, mfcc).detach().to('cpu')
-        gen_sample = gen_sample.squeeze(0).squeeze(0)
-        gen_sample = gen_sample.squeeze(1)
         gen_sample = gen_sample.numpy()
         gen_sample = self.vis_loader.dataset.denorm(gen_sample)
-        
-        sample_name = 'generated_sample_epoch_%03d' % epoch
-        save_dir = self.config.samples_dir / sample_name
-        np.save(save_dir, gen_sample)
+
+        gen_sample_batch_size = gen_sample.shape[0]
+        for sample_num in gen_sample_batch_size:
+            gen_sample_num = gen_sample[sample_num,:,:]
+
+            sample_name = f'gen_sample_{sample_num:03}'
+
+            save_dir = os.path.join(self.config.samples_dir, epoch)
+            if not os.path.exists(save_dir):
+                os.mkdir(save_dir)
+
+            save_path = os.path.join(save_dir, sample_name)
+            np.save(save_path, gen_sample_num)
 
 class LrwShapeTrainer(BaseTrainer):
     def __init__(self, config, train_loader,

### 2019-07-31 src/model/model/models.py
improved sampling saving

@@ -206,7 +206,7 @@ class MFCC_Shape_Gen(BaseModel):
         self.conv7 = nn.Conv2d(64, 32, kernel_size=(3,3), stride=(1,1))
         self.relu7 = nn.ReLU()
 
-        self.convt8 = nn.ConvTranspose1d(32, 4, kernel_size=3, stride=2)
+        self.convt8 = nn.ConvTranspose2d(32, 4, kernel_size=(1,3), stride=(1,2))
         self.tanh8 = nn.Tanh()
 
 
@@ -219,7 +219,6 @@ class MFCC_Shape_Gen(BaseModel):
         x = self.relu5(self.conv5(x))
         x = self.relu6(self.conv6(x))
         x = self.relu7(self.conv7(x))
-        x = x.squeeze(2)
         x = self.tanh8(self.convt8(x))
         return x
 
@@ -258,7 +257,6 @@ class MFCC_Shape_Critic(BaseModel):
 
     
     def forward(self, shapes, mfcc):
-
         # Expand shapes to same shape as mfcc
         height, width = mfcc.size(2), mfcc.size(3)
         ones = torch.ones(height, width, device=shapes.device)

### 2019-07-31 src/model/trainer/trainer.py

@@ -465,16 +465,17 @@ class MFCCShapeTrainer(BaseMultiTrainer):
     def save_sample(self, noise, mfcc, epoch):
 
         gen_sample = self.gen_model(noise, mfcc).detach().to('cpu')
+        gen_sample = gen_sample.squeeze(2)
         gen_sample = gen_sample.numpy()
         gen_sample = self.vis_loader.dataset.denorm(gen_sample)
 
         gen_sample_batch_size = gen_sample.shape[0]
-        for sample_num in gen_sample_batch_size:
+        for sample_num in range(gen_sample_batch_size):
             gen_sample_num = gen_sample[sample_num,:,:]
 
             sample_name = f'gen_sample_{sample_num:03}'
 
-            save_dir = os.path.join(self.config.samples_dir, epoch)
+            save_dir = os.path.join(self.config.samples_dir, str(epoch))
             if not os.path.exists(save_dir):
                 os.mkdir(save_dir)
 

### 2019-07-31 src/model/model/models.py
use hardtanh for clipping gen output. Linear activation clipped to [0,1]

@@ -207,7 +207,7 @@ class MFCC_Shape_Gen(BaseModel):
         self.relu7 = nn.ReLU()
 
         self.convt8 = nn.ConvTranspose2d(32, 4, kernel_size=(1,3), stride=(1,2))
-        self.tanh8 = nn.Tanh()
+        self.clip8 = nn.Handtanh(min_val=0, max_val=1)
 
 
     def forward(self, noise, mfcc):
@@ -219,7 +219,7 @@ class MFCC_Shape_Gen(BaseModel):
         x = self.relu5(self.conv5(x))
         x = self.relu6(self.conv6(x))
         x = self.relu7(self.conv7(x))
-        x = self.tanh8(self.convt8(x))
+        x = self.clip8(self.convt8(x))
         return x
 
 class MFCC_Shape_Critic(BaseModel):

### 2019-07-31 src/model/model/models.py
fixed hardtanh bug

@@ -207,7 +207,7 @@ class MFCC_Shape_Gen(BaseModel):
         self.relu7 = nn.ReLU()
 
         self.convt8 = nn.ConvTranspose2d(32, 4, kernel_size=(1,3), stride=(1,2))
-        self.clip8 = nn.Handtanh(min_val=0, max_val=1)
+        self.clip8 = nn.Hardtanh(min_val=0, max_val=1)
 
 
     def forward(self, noise, mfcc):

### 2019-07-31 src/model/model/models.py
another gen model. Without transpose

@@ -179,7 +179,7 @@ class old_MFCC_Shape_Gen(BaseModel):
         x = self.tanh10(self.conv10(x))
         return x
 
-class MFCC_Shape_Gen(BaseModel):
+class old2_MFCC_Shape_Gen(BaseModel):
     def __init__(self, z_dim, shapes_dim):
         super().__init__()
 
@@ -221,6 +221,65 @@ class MFCC_Shape_Gen(BaseModel):
         x = self.relu7(self.conv7(x))
         x = self.clip8(self.convt8(x))
         return x
+
+class MFCC_Shape_Gen(BaseModel):
+    def __init__(self, z_dim, shapes_dim):
+        super().__init__()
+
+        in_dim = z_dim + 1
+
+        self.conv1 = nn.Conv2d(in_dim, 32, kernel_size=(4,3), 
+                                           stride=(2,1), 
+                                           padding=(0,7))
+        self.relu1 = nn.ReLU()
+
+        self.conv2 = nn.Conv2d(32, 64, kernel_size=(4,3), stride=(2,1))
+        self.relu2 = nn.ReLU()
+
+        self.conv3 = nn.Conv2d(64, 128, kernel_size=(3,3), stride=(1,1))
+        self.relu3 = nn.ReLU()
+
+        self.conv4 = nn.Conv2d(128, 128, kernel_size=(3,3), stride=(1,1))
+        self.relu4 = nn.ReLU()
+
+        self.conv5 = nn.Conv2d(128, 128, kernel_size=(3,3), stride=(1,1))
+        self.relu5 = nn.ReLU()
+
+        self.conv6 = nn.Conv2d(128, 64, kernel_size=(3,3), stride=(1,1))
+        self.relu6 = nn.ReLU()
+
+        self.conv7 = nn.Conv2d(64, 4, kernel_size=(3,3), stride=(1,1))
+        self.sig7 = nn.Sigmoid()
+
+
+    def forward(self, noise, mfcc):
+        x = torch.cat((noise, mfcc), dim=1)
+        x = self.relu1(self.conv1(x))
+        x = self.relu2(self.conv2(x))
+        x = self.relu3(self.conv3(x))
+        x = self.relu4(self.conv4(x))
+        x = self.relu5(self.conv5(x))
+        x = self.relu6(self.conv6(x))
+        x = self.sig7(self.conv7(x))
+        return x
+
[…]

### 2019-08-01 src/model/datasets/datasets.py
improved normalisation of dataset and sample saving

@@ -109,8 +109,8 @@ class WavBlendshapesDataset(Dataset):
 
         self.stats = {'mfcc_min': np.inf,
                       'mfcc_max': -np.inf,
-                      'shape_min': np.inf,
-                      'shape_max': -np.inf}
+                      'shape_min': [np.inf] * self.n_shapes,
+                      'shape_max': [-np.inf] * self.n_shapes}
 
         self._collect_stats()
 
@@ -118,7 +118,7 @@ class WavBlendshapesDataset(Dataset):
         return len(self.wav_list)
     
     def __getitem__(self, idx):
-        mfcc, shape_param = self._get_data_pair(idx)
+        mfcc, shape_param, item_name = self._get_data_pair(idx)
 
         sample = {'mfcc': mfcc.astype(np.float32), 
                   'shape_param': shape_param.astype(np.float32)}
@@ -127,6 +127,8 @@ class WavBlendshapesDataset(Dataset):
 
         if self.transform:
             sample = self.transform(sample)
+        
+        sample['item_name'] = item_name
         
         return sample
     
@@ -147,7 +149,7 @@ class WavBlendshapesDataset(Dataset):
         crop_range = range(self.n_shapes, shape_param.shape[0])
         shape_param = np.delete(shape_param, crop_range, axis=0)
 
-        return mfcc, shape_param
+        return mfcc, shape_param, item_name
 
     def _collect_stats(self):
         """
@@ -159,24 +161,31 @@ class WavBlendshapesDataset(Dataset):
         except:
             print("Collecting dataset statistics...\n")
             for idx in tqdm(range(self.__len__())):
-                mfcc, shape_param = self._get_data_pair(idx)
+                mfcc, shape_param, _item_name = self._get_data_pair(idx)
 
                 idx_mfcc_min = mfcc.min()
                 idx_mfcc_max = mfcc.max()
-                idx_shape_min = shape_param.min()
-                idx_shape_max = shape_param.max()
+                idx_shape_min = [i for i in range(self.n_shapes)]
+                idx_shape_max = [i for i in range(self.n_shapes)]
+                for n in range(self.n_shapes):
+                    idx_shape_min[n] = shape_param[n,:].min()
+                    idx_shape_max[n] = shape_param[n,:].max()
 
                 if idx_mfcc_min < self.stats['mfcc_min']: 
                     self.stats['mfcc_min'] = idx_mfcc_min
                 if idx_mfcc_max > self.stats['mfcc_max']: 
[…]

### 2019-08-01 src/model/trainer/trainer.py

@@ -416,6 +416,7 @@ class MFCCShapeTrainer(BaseMultiTrainer):
         fixed_noise = torch.randn(self.vis_batch_size, self.z_dim, 
                                   height, width)
         fixed_noise = fixed_noise.to(self.device)
+        fixed_item_names = fixed_sample['item_name']
 
         for epoch in range(1, self.epochs+1):
             total_disc_loss = 0
@@ -458,22 +459,19 @@ class MFCCShapeTrainer(BaseMultiTrainer):
             self.logger.info('Critic Loss: {} '
                              'Gen Loss: {}'.format(disc_loss, gen_loss))
 
-            self.save_sample(fixed_noise, fixed_mfcc, epoch)
+            self.save_sample(fixed_noise, fixed_mfcc, fixed_item_names, epoch)
             if epoch % self.save_period == 0:
                 self._save_checkpoint(epoch)
     
-    def save_sample(self, noise, mfcc, epoch):
+    def save_sample(self, noise, mfcc, sample_names, epoch):
 
         gen_sample = self.gen_model(noise, mfcc).detach().to('cpu')
         gen_sample = gen_sample.squeeze(2)
         gen_sample = gen_sample.numpy()
         gen_sample = self.vis_loader.dataset.denorm(gen_sample)
 
-        gen_sample_batch_size = gen_sample.shape[0]
-        for sample_num in range(gen_sample_batch_size):
+        for sample_num, sample_name in enumerate(sample_names):
             gen_sample_num = gen_sample[sample_num,:,:]
-
-            sample_name = f'gen_sample_{sample_num:03}'
 
             save_dir = os.path.join(self.config.samples_dir, str(epoch))
             if not os.path.exists(save_dir):

### 2019-08-02 src/model/datasets/datasets.py
revert back to fully normalised shape params

@@ -109,8 +109,8 @@ class WavBlendshapesDataset(Dataset):
 
         self.stats = {'mfcc_min': np.inf,
                       'mfcc_max': -np.inf,
-                      'shape_min': [np.inf] * self.n_shapes,
-                      'shape_max': [-np.inf] * self.n_shapes}
+                      'shape_min': np.inf,
+                      'shape_max': -np.inf}
 
         self._collect_stats()
 
@@ -165,28 +165,22 @@ class WavBlendshapesDataset(Dataset):
 
                 idx_mfcc_min = mfcc.min()
                 idx_mfcc_max = mfcc.max()
-                idx_shape_min = [i for i in range(self.n_shapes)]
-                idx_shape_max = [i for i in range(self.n_shapes)]
-                for n in range(self.n_shapes):
-                    idx_shape_min[n] = shape_param[n,:].min()
-                    idx_shape_max[n] = shape_param[n,:].max()
+                idx_shape_min = shape_param.min()
+                idx_shape_max = shape_param.max()
 
                 if idx_mfcc_min < self.stats['mfcc_min']: 
                     self.stats['mfcc_min'] = idx_mfcc_min
                 if idx_mfcc_max > self.stats['mfcc_max']: 
                     self.stats['mfcc_max'] = idx_mfcc_max
 
-                for n in range(self.n_shapes):
-                    if idx_shape_min[n] < self.stats['shape_min'][n]: 
-                        self.stats['shape_min'][n] = idx_shape_min[n]
-                    if idx_shape_max[n] > self.stats['shape_max'][n]: 
-                        self.stats['shape_max'][n] = idx_shape_max[n]
+                if idx_shape_min < self.stats['shape_min']: 
+                    self.stats['shape_min'] = idx_shape_min
+                if idx_shape_max > self.stats['shape_max']: 
+                    self.stats['shape_max'] = idx_shape_max
 
             with open('data/lrw_audio_stats.pkl', 'wb') as f:
                 pickle.dump(self.stats, f, pickle.HIGHEST_PROTOCOL)
             
-            print(self.stats)
-
     def _mfcc(self, audio_data, sample_rate, n_mfcc=50):
         """
         Returns the mfcc of an audio signal.
@@ -229,12 +223,9 @@ class WavBlendshapesDataset(Dataset):
                     self.stats['mfcc_min'], 
                     self.stats['mfcc_max'])
         
-        min_vals = self.stats['shape_min']
-        max_vals = self.stats['shape_max']
-        for idx in range(self.n_shapes):
[…]

### 2019-08-02 src/model/datasets/datasets.py
LRW classifier added

@@ -63,7 +63,7 @@ class LrwBlendshapesDataset(Dataset):
         """
         Finds max and min values for blendshape params for dataset.
         """
-        print("Collecting dataset statistics...\n")
+        print("Collecting dataset statistics LRW...\n")
         for idx in tqdm(range(self.__len__())):
             shape_param, _label = self._get_sample(idx)
 
@@ -414,9 +414,8 @@ class LrwShapesToTensor(object):
 
         n_labels = 500
         label = torch.from_numpy(label)
-        #label = torch.nn.functional.one_hot(label, n_labels)
 
-        shape_params = torch.from_numpy(shape_params).unsqueeze(0)
+        shape_params = torch.from_numpy(shape_params)#.unsqueeze(0)
 
         return {'label': label, 'shape_params': shape_params}
 

### 2019-08-02 src/model/model/models.py

@@ -320,54 +320,53 @@ class Lrw_Shape_Classifier(BaseModel):
     def __init__(self):
         super().__init__()
 
-        self.conv1 = nn.Conv2d(1, 16, kernel_size=2)
+        self.conv1 = nn.Conv1d(4, 16, kernel_size=3)
+        self.bn1 = nn.BatchNorm1d(16)
         self.relu1 = nn.ReLU()
 
-        self.conv2 = nn.Conv2d(16, 32, kernel_size=2)
+        self.conv2 = nn.Conv1d(16, 32, kernel_size=3)
+        self.bn2 = nn.BatchNorm1d(32)
         self.relu2 = nn.ReLU()
 
-        self.conv3 = nn.Conv2d(32, 64, kernel_size=2)
+        self.conv3 = nn.Conv1d(32, 64, kernel_size=3)
+        self.bn3 = nn.BatchNorm1d(64)
         self.relu3 = nn.ReLU()
+        self.max_pool3 = nn.MaxPool1d(kernel_size=3, padding=1, stride=2)
 
-        self.conv4 = nn.Conv1d(64, 128, kernel_size=6)
+        self.conv4 = nn.Conv1d(64, 128, kernel_size=3)
+        self.bn4 = nn.BatchNorm1d(128)
         self.relu4 = nn.ReLU()
 
-        self.conv5 = nn.Conv1d(128, 256, kernel_size=6)
+        self.conv5 = nn.Conv1d(128, 256, kernel_size=3)
+        self.bn5 = nn.BatchNorm1d(256)
         self.relu5 = nn.ReLU()
+        self.max_pool5 = nn.MaxPool1d(kernel_size=3, padding=1, stride=2)
 
-        self.conv6 = nn.Conv1d(256, 256, kernel_size=6)
+        self.conv6 = nn.Conv1d(256, 128, kernel_size=3)
         self.relu6 = nn.ReLU()
 
-        self.conv7 = nn.Conv1d(256, 256, kernel_size=6)
-        self.relu7 = nn.ReLU()
-
-        self.conv8 = nn.Conv1d(256, 256, kernel_size=6)
-        self.relu8 = nn.ReLU()
-
-        self.conv9 = nn.Conv1d(256, 256, kernel_size=6)
-        self.relu9 = nn.ReLU()
-
-        self.linear10 = nn.Linear(2560, 2)
-        self.softmax10 = nn.functional.softmax
-
[…]

### 2019-08-02 src/model/train.py

@@ -12,7 +12,6 @@ def gan_main(config):
     logger = config.get_logger('train')
 
     data_loader = config.get('data_loader', data_loaders)
-    #vis_loader = data_loader.val_split()
     vis_loader = config.get('vis_loader', data_loaders)
 
     disc_model = config.get('discriminator,arch', models)
@@ -64,8 +63,8 @@ def classify_main(config):
 if __name__ == "__main__":
     fix_seed(0)
 
-    config = GetConfig('./config/mfcc_shape_gan/config.json')
-    gan_main(config)
+    #config = GetConfig('./config/mfcc_shape_gan/config.json')
+    #gan_main(config)
 
-    #config = GetConfig('./config/lrw_shape_classifier/config.json')
-    #classify_main(config)
+    config = GetConfig('./config/lrw_shape_classifier/config.json')
+    classify_main(config)

### 2019-08-02 src/model/model/models.py
added dropout to lrw classifer, improved logging

@@ -323,27 +323,33 @@ class Lrw_Shape_Classifier(BaseModel):
         self.conv1 = nn.Conv1d(4, 16, kernel_size=3)
         self.bn1 = nn.BatchNorm1d(16)
         self.relu1 = nn.ReLU()
+        self.dropout1 = nn.Dropout(0.1)
 
         self.conv2 = nn.Conv1d(16, 32, kernel_size=3)
         self.bn2 = nn.BatchNorm1d(32)
         self.relu2 = nn.ReLU()
+        self.dropout2 = nn.Dropout(0.1)
 
         self.conv3 = nn.Conv1d(32, 64, kernel_size=3)
         self.bn3 = nn.BatchNorm1d(64)
         self.relu3 = nn.ReLU()
         self.max_pool3 = nn.MaxPool1d(kernel_size=3, padding=1, stride=2)
+        self.dropout3 = nn.Dropout(0.1)
 
         self.conv4 = nn.Conv1d(64, 128, kernel_size=3)
         self.bn4 = nn.BatchNorm1d(128)
         self.relu4 = nn.ReLU()
+        self.dropout4 = nn.Dropout(0.1)
 
         self.conv5 = nn.Conv1d(128, 256, kernel_size=3)
         self.bn5 = nn.BatchNorm1d(256)
         self.relu5 = nn.ReLU()
         self.max_pool5 = nn.MaxPool1d(kernel_size=3, padding=1, stride=2)
+        self.dropout5 = nn.Dropout(0.1)
 
         self.conv6 = nn.Conv1d(256, 128, kernel_size=3)
         self.relu6 = nn.ReLU()
+        self.dropout6 = nn.Dropout(0.1)
 
         self.lin7 = nn.Linear(768, 500)
         self.softmax7 = nn.Softmax()
@@ -356,16 +362,16 @@ class Lrw_Shape_Classifier(BaseModel):
         """
         batch_size = shapes.size(0)
 
-        x = self.relu1(self.bn1(self.conv1(shapes)))
-        x = self.relu2(self.bn2(self.conv2(x)))
+        x = self.dropout1(self.relu1(self.bn1(self.conv1(shapes))))
+        x = self.dropout2(self.relu2(self.bn2(self.conv2(x))))
         x = self.relu3(self.bn3(self.conv3(x)))
-        x = self.max_pool3(x)
+        x = self.dropout3(self.max_pool3(x))
 
-        x = self.relu4(self.bn4(self.conv4(x)))
[…]

### 2019-08-02 src/model/trainer/trainer.py

@@ -503,7 +503,7 @@ class LrwShapeTrainer(BaseTrainer):
     
     def train(self):
         for epoch in range(1, self.epochs+1):
-            log = self._training_epoch(epoch)
+            self._training_epoch(epoch)
 
             if epoch % self.save_period == 0:
                 self._save_checkpoint(epoch)
@@ -533,27 +533,25 @@ class LrwShapeTrainer(BaseTrainer):
             self.optimizer.step()
 
             self.writer.set_step((epoch-1) * self.len_train_epoch + batch_idx)
-            self.writer.add_scalar('loss', loss)
+            self.writer.add_scalar('train/loss', loss)
 
             acc = correct / ((batch_idx+1) * self.batch_size)
-            #self.writer.set_step((epoch-1) * self.len_train_epoch + batch_idx)
-            self.writer.add_scalar('accuracy', loss)
+            self.writer.add_scalar('train/accuracy', acc)
 
             total_loss += loss
 
             if batch_idx % self.log_step == 0:
+                mean_loss = total_loss / ((batch_idx+1) * self.batch_size)
                 self.logger.info(
                     'Train Epoch: {} '
                     'Batch: {} '
                     'Loss: {:.6f} '
-                    'Accuracy: {:.6f}'.format(epoch, batch_idx, loss, acc))
+                    'Accuracy: {:.6f}'.format(epoch, batch_idx, mean_loss, acc))
 
         train_loss = total_loss / self.len_train_epoch
 
         if self.val_step:
-            val_loss = self._val_epoch(epoch)
-        
-        return {'train_loss': train_loss, 'val_loss': val_loss}
+            self._val_epoch(epoch)
 
     def _val_epoch(self, epoch):
         total_loss = 0
@@ -578,9 +576,15 @@ class LrwShapeTrainer(BaseTrainer):
                 acc = correct / ((batch_idx+1) * self.batch_size)
 
                 if batch_idx % self.log_step == 0:
-                    print('Val Epoch: {} Batch: {} '
-                          'Loss: {:.6f} Accuracy: {:.6f}'.format(
-                              epoch, batch_idx, loss, acc))
+                    mean_loss = total_loss / ((batch_idx+1) * self.batch_size)
+                    self.logger.info(
+                        'Val Epoch: {} '
+                        'Batch: {} '
+                        'Loss: {:.6f} '
+                        'Accuracy: {:.6f}'.format(
+                            epoch, batch_idx, mean_loss, acc))
 
[…]

### 2019-08-02 src/model/trainer/trainer.py
fixed validation bug

@@ -558,7 +558,7 @@ class LrwShapeTrainer(BaseTrainer):
         correct = 0
 
         with torch.no_grad():
-            for batch_idx, sample in enumerate(self.train_loader):
+            for batch_idx, sample in enumerate(self.val_loader):
                 step = (epoch-1) * self.len_train_epoch + batch_idx
 
                 label = sample['label']

### 2019-08-02 src/model/model/models.py
increase dropout to reduce overfitting

@@ -323,33 +323,33 @@ class Lrw_Shape_Classifier(BaseModel):
         self.conv1 = nn.Conv1d(4, 16, kernel_size=3)
         self.bn1 = nn.BatchNorm1d(16)
         self.relu1 = nn.ReLU()
-        self.dropout1 = nn.Dropout(0.1)
+        self.dropout1 = nn.Dropout(0.2)
 
         self.conv2 = nn.Conv1d(16, 32, kernel_size=3)
         self.bn2 = nn.BatchNorm1d(32)
         self.relu2 = nn.ReLU()
-        self.dropout2 = nn.Dropout(0.1)
+        self.dropout2 = nn.Dropout(0.2)
 
         self.conv3 = nn.Conv1d(32, 64, kernel_size=3)
         self.bn3 = nn.BatchNorm1d(64)
         self.relu3 = nn.ReLU()
         self.max_pool3 = nn.MaxPool1d(kernel_size=3, padding=1, stride=2)
-        self.dropout3 = nn.Dropout(0.1)
+        self.dropout3 = nn.Dropout(0.2)
 
         self.conv4 = nn.Conv1d(64, 128, kernel_size=3)
         self.bn4 = nn.BatchNorm1d(128)
         self.relu4 = nn.ReLU()
-        self.dropout4 = nn.Dropout(0.1)
+        self.dropout4 = nn.Dropout(0.2)
 
         self.conv5 = nn.Conv1d(128, 256, kernel_size=3)
         self.bn5 = nn.BatchNorm1d(256)
         self.relu5 = nn.ReLU()
         self.max_pool5 = nn.MaxPool1d(kernel_size=3, padding=1, stride=2)
-        self.dropout5 = nn.Dropout(0.1)
+        self.dropout5 = nn.Dropout(0.2)
 
         self.conv6 = nn.Conv1d(256, 128, kernel_size=3)
         self.relu6 = nn.ReLU()
-        self.dropout6 = nn.Dropout(0.1)
+        self.dropout6 = nn.Dropout(0.2)
 
         self.lin7 = nn.Linear(768, 500)
         self.softmax7 = nn.Softmax()
